### 2014-10-15 AdamRebuttal.txt
stuff

Reviewer A:

We want to emphasize why the result about sequential consistency is important.
Here we use "sequential consistency" as a name for the LTS in Figure 3, which
actually completely specifies the observable behavior of a full multiprocessor
system.  Thus, by proving a relationship with that LTS, we are proving what is
perhaps the most natural version of "full functional correctness" in this
domain.

Regarding store atomicity and weak memory models: we define store atomicity as
a property of main memory only, which is quite complex in its own right, with
hierarchies of caches, etc.  There are additional cache-like elements attached
to the individual processors, called by names like "store buffer".  If we
consider these extra buffers as part of the memory, then you're certainly right
that weak memory models don't guarantee atomic stores. However, weak-memory
systems can be built on top of the same store-atomic main memory interface we
give in the paper; it's only the processor-local store buffers that introduce
weaker behavior.

We agree, though, that we should include less speculation on adaptation to weak
memory; and thanks for the bug reports on our discussion of related work.

Reviewer B:

We agree that the underlying LTS proof techniques are not new.  We view our
contribution as more of demonstrating a new application in an area that badly
needs techniques at this level of modularity.

For both Reviewers B and C:

Sorry for not explaining more of our rationale for considering these LTSes as
practical hardware descriptions.  The cache-coherence algorithm, for instance,
is directly transliterated out of [fill in self-contained paper references
here].  We would like to have an automatic translation from Bluespec code, but
we think the two abstraction levels are close enough that it is reasonable to
draw conclusions from our manual translations.

Reviewer C:

Our main theorem uses termination as an arbitrary property that is sufficient
to make the theorem interesting but still easily stated. That's for the same
reason that many problems in computability theory are equivalent to the halting
problem in suitable senses.

Reviewer D:

We apologize for not getting across the sense of the main theorem. Our
optimized architecture actually does fulfill the SC spec, without any side
conditions on programs!  The explanation of why is not trivial, and the proof
carried out throughout the paper justifies the claim.

We also view simplicity of the proof outline (at the level given in the paper)
as a feature, not a bug, of the approach. The actual proof in Coq is much
longer and more detailed (about 10,000 lines).