### 2014-10-15 ArvindRebuttalLatest.txt

We are thankful to the reviewers for their valuable and in-depth comments. We
hope the following clarifications would make them more enthusiastic about
advocating its publication in POPL.

1. Store-atomicity is ubiquitous: Even though systems implement memory models
weaker than sequential consistency (SC), real multiprocessor systems can be
decomposed into multiple parts, one of which is an atomic-memory. In the
simplest form, atomic memory is just a monolithic multiported global memory
(GM), but more commonly, it is implemented using a hierarchy of coherent
caches. Weak memory model behaviors can be understood in terms of local storage
(LS) between each processor and GM. If the processor memory interface is drawn
such that LS is part of GM, then GM will not behave like an atomic store. By
making LS part of the processor model, we assert that all the nuances of the
known weak memory models can be modeled as part of the processor description.
We address the two examples cited by Reviewer A at the end. The proof of
correctness of cache coherence is completely orthogonal to the semantics of the
processor side and hence the memory model issues.

2. Choice of SC: The processor description in Figure 7 implements SC. This
description will have to be modified for weaker memory models, but the
correctness criteria for ROB in Figure 6 will remain the same. For example, one
may not even issue a verification load or issue it sooner. We acknowledge that
more work needs to be done to formally model weaker memory models, including a
treatment of fences.

3. SC performance issues: In the current architecture literature, it's hotly
debated whether weaker memory models can provide better performance than
properly implemented SC. Our SC processor in Figure 7 has high performance
because it is not restricted to sustain only a fixed number of (speculative)
loads in flight. Furthermore, instructions dependent upon these loads can be
processed out of order. A weaker memory model cannot improve upon this
important aspect. The main restriction in our model is that a processor can
issue only one store or one verification load which may be improvable by a
weaker model.

4. LTS for describing and synthesizing hardware:  A Bluespec program is a
transition system and experience has shown that Bluespec descriptions can be
both concise and generate efficient hardware. The cache coherence protocol of
Figure 10 was initially implemented and synthesized into efficient hardware
from Bluespec (Khan et al ISPASS-2012, see also Dave et al Memocode 2006). A
Bluespec program can be transliterated into an LTS in Coq; currently, this is
done manually.

The processor LTS in Figure 7 has an abstract specification of the ROB and
branch predictor with a clearly stated correctness criteria. To derive the
actual hardware, the LTSes for these components have to be specified in much
greater detail. Our ISA is general enough to encompass any CISC ISA including
x86. However the complexity of the detailed LTSes for ROB etc will depend upon
the specific ISA.

5. Reusability of our Cache Coherence proof: MOSI, MOESI, Firefly etc. are very
similar to MSI protocol, though their transition systems will be different. The
way our proof is organized, we will have to prove Lemmas 5 and 6, and
Invariants 2, 3, and 4 for each of these protocols; the rest of the proof will
remain unchanged. In the current version, these lemmas and invariants account
for x% of the proof. Also all these lemmas deal with less than 4 states and
thus can be proven easily using model checking.

------- 
Examples For Reviewer A

For explaining Reviewer A's IRIW behavior, consider a system containing several
in-order cores, each core connected to what we call as ''non-coherent cache''
NCC. All NCCs are connected to a global atomic memory. An NCC can read any
location from memory and store it, and the core can use the data from NCC, if
present. But stores from the core always happen to the global memory, and the
NCC can discard the location that it has, anytime. So, in the example,
initially, NCC of core 4 reads in the location x. Now the sequence of
operations is
Core 1: Wx1 (into global memory)
Core 2: Rx (gets 1 from global memory)
        Ry (gets 0 from global memory, initial value)
Core 3: Wy1 (into global memory)
Core 4: Ry (gets 1 from global memory)
        Rx (gets 0 from NCC)

Similarly, the second example about store buffers can be explained using a
store buffer instead of the NCC, into which stores are written, and loads can
read from. The sequence for this to explain the behavior is:
Core 1: Wx1 (into store buffer, not to memory)
        Ry (read from memory 0)
Core 2: Wy1 (into store buffer, not to memory)
        Rx (read from memory 0)

This global memory is, in reality, implemented using cache coherent protocols.
The separation of coherent caches implementing atomic memory semantics and the
rest of the system is an important contribution of this paper; a version of
this was presented in Arvind (Memory Model = Instruction Reordering + Store
Atomicity, ISCA 2006), and this paper gives a formal proof of this separation,
while extending this idea to speculating systems.