### 2014-10-15 Acknowledgements.tex

%\subsection*{Acknowledgments}
{\footnotesize\noindent\textbf{Acknowledgments:} The parts of this work done
at MIT was supported by Quanta Research agreement dated 4/1/05 and
Angstrom Project under Grant No. HR0011-10-9-0009. The research at SRI
was supported by Defense Advanced Research Projects Agency (DARPA) and
Air Force Research Laboratory (AFRL), under contract FA8750-11-C-0249
and National Science Foundation under Grant No. CCF-1217498. The
views, opinions, and/or findings contained in this report are those of
the authors and should not be interpreted as representing the official
views or policies, either expressed or implied, of the Quanta
Research, the Defense Advanced Research Projects Agency, the United
States Air Force or the Department of Defense. We would like to thank
Richard Uhler for helping focus this paper in early drafts.% }
%
%{ We would like to thank Richard Uhler for helping focus this paper in
%  early drafts.  This work was supported by the Defense Advanced
%  Research Projects Agency (DARPA) and the United States Air Force,
%  under Contract No. FA8750-11-C-0249. Any opinions, findings and
%  conclusions or recommendations expressed in this material are those
%  of the authors and do not necessarily reflect the views of DARPA or
%  the United States Air Force.  }