### 2014-10-15 AtomicStore.tex

\section{Enabling Practical Memory Subsystems: Atomic Store}

The behavior of a memory system realizing different memory models or
interfacing with different levels speculation from the processors would be
dramatically different. For instance, a memory system realizing Total Store
Order allows significantly the more memory requests to be in flight than in a
sequential consistent system. As such one might expect that there should be
relatively little sharing between designs.

%% At a high-level, the physical interface of memory is fairly consistent
%% regardless of the memory model or instruction execution scheme. The processor
%% makes requests (loads and stores) and gets responses back from the
%% memory.

Memory subsystems are designed to allow parallel operation

The most natural interface for a memory subsystem is to expect the interface to

Memory needs caches, highly pipelined. All worries happen at a single address. No ordering guarantees between two different addresses. This is a natural representation because.

Clearly this model is not SC. However, it is highly atuned to high-performance results.