### 2014-10-15 Implementation.tex

\section{How to Build and Sequentially System using Atomic Store}

We can have a simple system using a lamport-like processor which will nopipelining of memory. Trivially SC with atomic store. This is obviously correct, but from a point of view of practice, we must pipeline this. We need to refine this model into a

(How to use a non-monolithic memory).

\section{How to build a monolithic processor}

A speculative Lamport system consists of a monolithic memory and other loads and stores. This is how this system works. Need to discuss

* need definitions of spec processor. (Must be complete and sound)

A full realistic implementation: