### 2014-10-15 Introduction.tex

\section{Introduction}
\label{sec:Introduction}

A modern high-performance, cache-coherent, distributed-memory hardware system is
inherently complex. Such a system implements memory models like sequential
consistency (SC)~\cite{lamport1979make}, Total Store Order (TSO), and
Relaxed Memory Order (RMO)~\cite{weaver1994sparc}, each
of which characterizes how loads and stores may be issued by a processor and
how the shared-memory component responds to these requests.  These systems by
their nature are highly concurrent and nondeterministic.

Hardware designers recognize the challenge of implementing such systems, to such an extent that they have
already become some of the most serious real-world adopters of formal methods.
Hardware verification is dominated by model-checking, where abstraction
techniques are used to reduce designs to finite state spaces, which can be
explored exhaustively.  There are limits to the construction of sound
abstractions, so verifications of, e.g., cache-coherence protocols have almost
always treated systems with concrete topologies, involving particular finite
numbers of caches and processors. For instance, explicit-state model checking tools
like Murphi~\cite{murphi} or TLC~\cite{tlc} are only able to handle small and
simplified systems, \eg{} a single-level cache hierarchy with three CPUs, two
addresses, and one bit of data. Symbolic model checking techniques have fared
slightly better; McMillan \etal{} have verified a 2-level MSI protocol based on
the Gigamax distributed multiprocessor using SMV~\cite{gigamax}. Optimizations
on these techniques (\eg{} partial order reduction~\cite{part}, symmetry
reduction~\cite{sym1, sym2}, compositional reasoning~\cite{comp}, \etc{}) scale
the approach further, but they are still unable to handle realistic
hierarchical protocols. We might hope to achieve higher
assurance and understanding of design ideas by verifying \emph{infinite families}
of hardware designs, which resist reduction to finite-state models.  A
potential related benefit is reducing the computational resources required
to formally verify a new design in the family, applying a more general theorem without
much new analysis.

A related benefit of a deductive approach is \emph{modularity}.  It seems to
be rare for hardware practitioners to verify full designs that include
distinct functional components, like processors and memory systems.  One
reason is the algorithmic overhead of exploring such large state spaces.
If we can instead give localized correctness theorems for different
subsystems, then we can compose these theorems in a black-box way to
produce full-system theorems.

The traditional notion of modularity in hardware is closely tied to
\emph{Finite State Machines} (FSMs), more specificially synchronous FSMs. The biggest
problem with using FSM abstraction is that it restricts significantly the kinds of
refinements that can be performed on a module. For example, if we change the
timing of an adder so that it takes 2 clock cycles instead of 1, the whole
system is likely to break. Even though big modules in hardware are often
designed with asynchronous handshaking protocols, current
hardware verification methodologies do not deal with timing refinements.
In this paper, we use the formalism of \emph{labeled transition systems} that
is well-known in process-calculus circles.  It turns out to be exactly the
right way of capturing contracts on interaction between hardware components,
enabling both modular refinement and verification.

%Proofs in this style are quite common in the software world.  The function
%preconditions and postconditions of Hoare logic provide a canonical
%example, and there have been many innovations in modular proof since Hoare's
%original work.  Somehow the hardware verification world has taken less
%advantage of modularity.  In this paper, we propose a new program in
%hardware verification, based on modular proofs using the formalism of
%\emph{labelled transition systems} that is well-known in process-calculus
%circles.  It turns out to be exactly the right way of capturing contracts on
%interaction between hardware components.

Computer architects have intuitions about the contracts of classic
subsystems.  However, these intuitions have not been formalized
in enough detail to enable rigorous proofs.  In carrying out that exercise,
we have also learned some things about architectural design that should
be of interest outside a formal-methods context.

Concretely, we provide a modularization of distributed shared-memory systems
that makes minimal assumptions about networks, provides an abstraction for
memory that allows high-performance implementation of any reasonable memory
model, and allows a full range of realistic performance optimizations in
processors without violating modularity. We want our modular system to permit
\emph{black-box refinement}, that is, nothing beyond the interface contract can
be assumed about other modules.

In each instance we describe a reference model and an implementation model for
a module using labeled transition systems and formally prove that the
implementation is correct (sound) with respect to our reference. For example,
our reference model for memory is monolithic, where each load or store request
generates a (conceptually) instantaneous response, but where requests may be processed
in any order. Such a model is intuitively \emph{store atomic} in the sense that
the effect of a store is visible to all subsequent loads. We then describe a
realistic cache-coherence protocol comprised of a hierarchy of caches, and
formally prove that this system can never produce
a behavior that is not allowed by the reference model. Similarly, we define a
reference model for a processor that executes one instruction at a time and
an implementation model that exploits speculation to increase parallelism,
and we show that the speculative processor implements the reference processor,
in a suitable sense.

In the end, we prove a theorem about an infinite family of hardware systems,
each system having an arbitrary number of speculative processors connected to
a memory with an arbitrary hierarchy of caches.  All our proofs are mechanized
in the Coq proof assistant, and they are modular in the sense of allowing
further optimization to either processor or memory without needing to touch
the proof of the other.  We characterize overall system behavior with a very
simple transition system formalizing \emph{sequential consistency}, and we
conjecture that the methodology generalizes to other popular memory models.

\begin{figure}
\centering
\includegraphics[scale=.43]{zoom}
\caption{Overall system}
\label{zoom}
\end{figure}

Figure \ref{zoom} sketches the sort of hardware system that we have verified.
We show the top-level system design at top right, and we zoom in on boxes standing for the
memory system and a single processor.

Memory is composed of a \emph{hierarchy of caches}, where each cache node (labeled
like ``L1,'' ``L2,'' etc.) communicates only with its neighbors in the graph.
Each processor has a dedicated L1 cache, from which we do our best to satisfy
all memory requests, to avoid the latency of a round-trip with main memory.
With an L1 cache miss, we may still find a hit in a parent cache below the
level of main memory, realizing a smaller but still significant speedup.
We have verified a \emph{directory-based} protocol for coordinating an arbitrary
tree of caches, where each node stores a conservative approximation of its
children's states.

A processor is decomposed into several components.  We have the normal
architectural state, such as values of registers.  Our proofs are generic over
\emph{a family of instruction set architectures}, with parameters for opcode sets and
functions for executing opcodes and decoding them from memory.  Other key
components are a \emph{branch predictor}, which guesses at the control-flow path
that a processor will follow, to facilitate speculation; and a
\emph{reorder buffer (ROB)}, which decides which instructions along that path to
try executing ahead of schedule.  Our proofs apply to an arbitrary branch predictor,
and they work for any reorder buffer satisfying a simple semantic condition.

Figure~\ref{proofs} gives the overall proof structure that we employ to verify
the system of Figure~\ref{zoom}. As we will explain later, $P_\text{so}$
represents the speculative out-of-order processor, $M_c$ the cache memory,
$M_m$ the simple memory, $P_\text{ref}$ a simple decoupled processor, and
SC the simple reference model having multiple processors executing each
instruction atomically thus implementing sequential consistency.  Notation
$A^n$ is for $n$ copies of system $A$ running in parallel.
$\sqsubseteq$ is a refinement operator, capturing a suitable notion of when
one system implements another.  We go into more detail on the symbols from
the figure in the rest of the paper.

Our current framework establishes theorems of the form ``if system $A$ has a run
with some particular observable behavior, then system $B$ also has a run with
the same behavior.''  In this sense, we say that $A$ correctly implements $B$.
Other important properties, such as \emph{deadlock freedom} for $A$ (which
might get stuck without producing any useful behavior), we leave for future
work.

\begin{figure}
\includegraphics[scale=.45]{proofs}
\caption{Overall proof structure}
\label{proofs}
\end{figure}

The main contributions of this paper are:

\begin{itemize}

\item The \emph{first machine-checked proof} of an implementation of sequential
  consistency involving speculative processors and cache-coherent distributed
  memory.

\item The \emph{first machine-checked proof} of an invalidation-based
  cache-coherence protocol for distributed memories comprised of arbitrary
  cache hierarchies. We believe such a memory is also the correct abstraction for
  implementing relaxed memory models.

\item A general \emph{modular verification methodology} applicable to hardware
  algorithms, based on \emph{labeled transition systems}.  We are able to verify
  the several components of Figure~\ref{zoom} against general formal interfaces,
  enabling mixing and matching of different realizations of each interface,
  without doing any new proofs that peek beneath abstraction boundaries.

\item We use semantics to \emph{elucidate folklore ideas in computer architecture}.
  For instance, we formalize a variety of processor speculation patterns with
  the idea of speculative loads followed by verification loads, and we capture
  a key requirement on memory systems as refinement of a simple transition
  system.

\end{itemize}

As a side effect, we %debunk a common misconception about the semantics of the
have shown a (rather surprising) architectural result about the semantics of
the interface between processor and memory system. One can establish a
sequentially consistent system even if the buffers between the memory system
and the processor do not preserve message order. We prove as much by
establishing simulation of sequential consistency without relying on an
ordering assumption.

\paragraph{Paper Organization:} In Section~\ref{sec:lts} we introduce
our flavor of the labeled transition systems formalism,
including a definition of trace refinement. In
Section~\ref{sec:sc} we give a simple formal model of sequential consistency,
which we decompose further for decoupled shared-memory systems in Section~\ref{sec:ref}.
The following sections refine the two main subcomponents from Figure~\ref{zoom}.
Section~\ref{sec:ooo} discusses definition and verification of a
speculative processor model, and Sections~\ref{sec:cc} and~\ref{sec:ccproof}
respectively define and prove our hierarchical cache-coherence protocol.
Finally, in Section~\ref{sec:finalresult} we show
the whole-system modular proof of our complex system, followed by
discussion of implications in Section~\ref{sec:implications}
and further discussion of related work and other conclusions in
Sections~\ref{relatedWork} and \ref{sec:conclusion}.

The Coq source code of our proofs is included as the non-anonymous
supplement to this paper.

%% Modularity has long been understood as a key property for effective design and
%% verification of complex systems, \eg\ distributed memory systems. For the
%% designer, it allows a separation of concerns increasing robustness by allowing
%% the behavior encapsulated by a modular boundary to be realized by multiple
%% implementation any of which may be safely dropped into the system. It also
%% allows a greater measure of parallelization of design improving development
%% time. Similarly verification is simplified as modular interface agreements
%% provide a set of natural lemmas to verify. This leads us towards a
%% decomposition of the whole system verification task into smaller more
%% manageable chunks.

%% Given its complexity, it is a practical necessity for distributed shared memory
%% systems to be built a strong idea of modularity. The standard modularization
%% (see Figure~\ref{fig:highleveldsm}) separates the ISA-level processors nodes
%% which are responsible for executing threads of computation, and a unified
%% shared memory subsystem which is responsible for ensuring that memory requests
%% it receives are handled correctly, \ie{} in line with the memory model which
%% dictates which store values a load may observe in the system. These processors
%% may then be refined for performance, while the monolithic memory module is
%% improved by introducing layers of caching and a coherence protocol to guarantee
%% memory consistency.

%% This high-level approaches works reasonably well for implementation; separate
%% design groups tackle can each task in isolation. However, when we consider
%% verification, the modularity enforced during the implementation has little
%% benefit and in practice we must resort to full system verification. The core of
%% this problem lies in the fact that the simple high-level decomposition of
%% memory and processors does not fully take into account the space of speculation
%% techniques used in the processor.

%% All practical processor instruction set architectures (ISAs) are fairly
%% sequential; the result of executing an instruction may affect which next
%% instruction is executed. As a result for performance some sort of prediction is
%% used to allow instructions to be speculatively started and executed before the
%% previous instructions have completed and the values need to execute are known.
%% Additional bookkeeping is used to determine when our guess is wrong and
%% false-path instructions are undone leaving no affect on the architectural
%% state. As precise exception handling is a requirement these speculative
%% instructions are committed, \ie{} expose their effects of execution to the
%% whole system, in sequence.

%% Though most forms of speculation (\eg{} branch prediction, instruction prefetch
%% from pipelining, value prediction) are often considered as being orthogonal to
%% memory issues, speculation changes the sequence of memory requests the
%% processor sends to the memory. This may be merely handling requests out of
%% order, \eg{} issuing a load request before a logically earlier one has issued
%% (a common occurrence due to data dependencies in out-of-order processors). We
%% may also allow incorrect memory operations to be issued, \eg{} executing a load
%% from a false path. It is worth noting that while such load operations are
%% possible to speculate, we cannot speculatively issues store operations to the
%% memory as they become exposed to the whole system and speculation mechanisms
%% are isolated to a single processor. As a consequence of this, our memory does
%% not know what order requests it is sent should happen. Thus the memory in
%% isolation cannot directly realize the memory model and in actually implements a
%% much weaker model leaving the full enforcement of the memory model to the
%% processor which can enforce this by restricting when it issues requests.

%% The goal of this work is to introduce an interface enabling true modular
%% implementation and verification of shared memory systems. That is, we should be
%% able to describe a modular decomposition where we may correct replace either
%% the processor nodes, the memory with locally verified alternatives without
%% changing the correctness of our whole system proof of correctness or
%% restricting our ability to describe shared memory systems. As part of this we
%% will describe the processor interface in a way which allows the full range of speculation

%% We do so by showing the construction and initial refinements of a sequentially
%% consistent shared memory system. We formally defining a minimal cache coherent
%% speculation-friendly memory interface (\emph{Store Atomicity}), and use this as
%% the interface for our memory subsystem. Using this we specify the notion of
%% correct for a processor and prove that our system is correct assuming the
%% correct implementation of these two objects.  We then construct both a naive
%% reference processor and memory subsystem and provide proofs of correctness in
%% Coq. We then refine both components into more realistic implementations
%% including taking our memory subsystem to hardware-level description of a
%% arbitrary-level hierarchical cache coherence implementation and verify these
%% module. We show that the cross products of possible implementations are correct
%% and their proof automatically from the modularly.

%%  LocalWords:  Modularity modularity