### 2014-10-15 SequentialConsistency.tex

\section{Defining Sequential Consistency}

To dicsuss the modularization of our system we must first fully define what we
mean by sequential consistency. Informally, sequential consistency states that
the memory operations from all processors in the system are evaluated
atomically with regards to each other by the memory and are evaluated in order
locally but with no ordering requirements between operations from different
processors.

Formally Sequential Consistency can be described as:

\xxx{SC Definition HERE}.

\xxx{TOTAL ORDERING CONDITION}

Total ordering is not generally stated.