### 2014-10-15 Abstract.tex
all the changes

We provide a framework for modular verification of multi-processor hardware
systems with respect to memory model specification. In this framework, we make
use of a key observation that eventhough different memory models have different
overall system behaviors, they all share a coherent cache hierarchy. We prove
that this coherent cache hierarchy implements the \emph{store atomicity}
property, which makes it behave like an atomic instantaneous memory with
respect to the rest of the system. We further modularize the proof by defining
what is meant for a processor to be correct independently of the memory system.
As a concrete instance of our framework, we walk through a modular proof of
high-performance multi-processor implementation of Lamport's Sequential
Consistency model. The processors employ \emph{speculative execution} to
generate many concurrent loads to the memory sub-system, while obeying our
correctness criteria. We offer the first formal proof that when an atomic
memory is connected to $n$ speculative processors, the system as a whole is
sequentially consistent.  Additionally, the cache-coherent memory we present in
this paper is the first fully verified system that supports arbitrary cache
hierarchies. Our framework involves describing all our components using labeled
transition systems and we machine-verify all our proofs using the Coq proof
assistant.