### 2014-10-15 TransitionsUseful.tex

\section{Labeled Transition Systems}\label{sec:lts}

We make extensive use of the general theory of labeled transition
systems, a semantics approach especially relevant to communicating
concurrent systems.  As we are formalizing processors for
Turing-complete machine languages, it is challenging to prove that a
system preserves almost any aspect of processor behavior from a model
like SC.  To focus our theorems, we pick the time-honored property of
\emph{termination}.  An optimized system should terminate or diverge
iff the reference system could also terminate or diverge,
respectively.  Our basic definitions of transition systems build in
special treatment of halting, so that we need not mention it
explicitly in most contexts to come.

\begin{defn}
A \textbf{labeled transition system (LTS)} is a ternary relation, over
$\mathcal S^H \times \mathcal L^\epsilon \times \mathcal S^H$, for some sets
$\mathcal S$ of states and $\mathcal L$ of labels. We usually do not mention
these sets explicitly, as they tend to be clear from context. We write
$X^\epsilon$ for lifting of a set $X$ to have an extra ``empty'' element
$\epsilon$ (like an \texttt{option} type in ML). We write $X^H$ for lifting of
a set $X$ to have an extra ``halt'' element $H$. We also implicitly consider
each LTS to be associated with an \emph{initial} state in $\mathcal S$.
\end{defn}

For LTS $A$, we write $\io{A}{s}{s'}{\ell}$ as shorthand for $(s, \ell, s') \in
A$, and we write $A_0$ for $A$'s initial state. The intuition is that $A$ is
one process within a concurrent system. The label $\ell$ from set $\mathcal L$
of labels is produced when $A$ participates in some IO exchange with another
process; otherwise it is an empty or ``silent'' label $\epsilon$. As a
shorthand, we sometimes omit labels for $\epsilon$ steps.

\subsection{Basic Constructions on LTSes}

From an LTS representing single-step system evolution, we often want to build
an LTS capturing arbitrary-length evolutions.

\begin{defn}
The \textbf{transitive-reflexive closure} of $A$, written $A^{*}$, is a derived
LTS.  Where $A$'s states and labels are $\mathcal S$ and $\mathcal L$, the
states of $A^{*}$ are $\mathcal S$, and the labels are $\mathcal L^{*}$, or
\emph{sequences} of labels from the original system.  $A^{*}$ steps from $s$ to
$s'$ when there exist zero or more transitions in $A$ that move from $s$ to
$s'$.  The label of this transition is the \emph{concatenation} of all labels
generated in $A$, where the empty or ``silent'' label $\epsilon$ is treated as
an identity element for concatenation.
\end{defn}

We also want to compose $n$ copies of an LTS together, with no explicit
communication between them.  We apply this construction later to lift a
single-CPU system to a multi-CPU system.

\begin{defn}
The \textbf{$n$-repetition} of $A$, written $\smulti{n}{A}$, is a derived LTS.
Where $A$'s states and labels are $\mathcal S$ and $\mathcal L$, the states of
$\smulti{n}{A}$ are $\mathcal S^n$, and the labels are $[1, n] \times \mathcal
L$, or pairs that \emph{tag} labels with which component system generated them.
These labels are generated only when the component system generates a label.
The whole system halts whenever one of the components halts.  We define the
transition relation with the following inference rules. We treat $n$-tuples as
functions keyed on numeric positions, and we write $\theta[i := v]$ for
updating $n$-tuple $\theta$ to overwrite the $i$th position with value $v$, and
we treat any label $(i, \epsilon)$ as $\epsilon$.
$$\inference[i]
{\theta(i) = a & \io{$A$}{a}{a'}{\ell} & a' \neq H}
{\io{$\smulti{n}{A}$}{\theta}{\theta[i:=a']}{i,\ell}}
\quad \inference[H]
{\io{$A$}{\theta(i)}{H}{\ell}}
{\io{$\smulti{n}{A}$}{\theta}{H}{i,\ell}}
$$
\end{defn}

Eventually, we need processes to be able to communicate with each
other, which we formalize via the $+$ composition operator. It
connects same-label transitions in the two systems, treating the label
as a cooperative communication event that may now be hidden from the
outside world, as an $\epsilon$ label.

\begin{defn}\label{plus}
Where $A$ and $B$ are two LTSes sharing labels set $\mathcal L$,
and with state sets $\mathcal S_A$ and
$\mathcal S_B$ respectively, the \textbf{communicating composition} $A + B$ is
a new LTS with states $\mathcal S_A \times \mathcal S_B$ and an empty label set,
defined as follows:
\vspace{-.2in}\begin{center}
$$\inference[$A$]
{\io{$A$}{a}{a'}{} & a' \neq H}
{\io{$A + B$}{a, b}{a', b}{}}
\quad \inference[$B$]
{\io{$B$}{b}{b'}{} & b' \neq H}
{\io{$A + B$}{a, b}{a, b'}{}}$$
$$\inference[H$_A$]
{\io{$A$}{a}{H}{}}
{\io{$A + B$}{a, b}{H}{}}
\quad \inference[H$_B$]
{\io{$B$}{b}{H}{}}
{\io{$A + B$}{a, b}{H}{}}$$
$$\inference[Join]
{\io{$A$}{a}{a'}{\ell} & \io{$B$}{b}{b'}{\ell} & a', b' \neq H}
{\io{$A + B$}{a, b}{a', b'}{}}$$
\end{center}
\end{defn}

\subsection{Refinement Between LTSes}

We need a notion of when one LTS ``implements'' another.  Intuitively,
transition labels and halting are all that the outside world
can observe. Two systems that produce identical labels and termination behavior
under all circumstances can be considered as safe substitutes for
one another. We will only need an asymmetrical notion of compatibility:

\begin{defn}\label{refines}
For some label domain $\mathcal L$, let $f : \mathcal L \to \mathcal L^\epsilon$ 
be a function that is able to replace labels with alternative
labels, or erase them altogether. Let LTSes $A$ and $B$ have the same label
sets $\mathcal L$. We say that
\textbf{$A$ trace-refines $B$ w.r.t. $f$}, or $A \sqsubseteq_f B$, if:
\begin{multline*}
\forall s_A, \eta. \; \io{$A^{*}$}{A_0}{s_A}{\eta} \Rightarrow \exists s_B. \;
\io{$B^{*}$}{B_0}{s_B}{f(\eta)} \; \wedge \\
(s_A = H \Leftrightarrow s_B = H)
\end{multline*}
Each label in the trace is replaced by the mapping of $f$ on it, and
labels mapped to $\epsilon$ by $f$ are dropped. $f$ is overloaded
to denote that multi-label version when applied to $\eta$.
\end{defn}

As a shorthand, we write $A \sqsubseteq B$ for $A \sqsubseteq_\mathsf{id} B$,
for $\mathsf{id}$ an identity function, forcing traces in the two systems to
match exactly. Under this notion of identical traces, we say that $A$
is sound w.r.t. $B$.

\subsection{A Few Useful Lemmas}

The reflexivity and transitivity properties of refinement are crucial to help us
structure modular proofs.

\begin{theorem}\label{transitive}
$\sqsubseteq$ is reflexive and transitive.
\end{theorem}

Another crucial proof step is to lift a single-process result to the $n$-way
replication of that process.

\begin{theorem}\label{liftn}
If $A \sqsubseteq_f B$, then $\smulti{n}{A} \sqsubseteq_{f^n} \smulti{n}{B}$, where
$f^n$ is $f$ lifted appropriately to deal with indices ($f^n(i, \ell) =
(i, \ell')$ when $f(\ell) = \ell'$, and $f^n(i, \ell) = \epsilon$ when
$f(\ell) = \epsilon$).
\end{theorem}

We will also need a result applying to communicating compositions.

\begin{theorem}\label{liftplus}
If $A \sqsubseteq_f A'$ and $B \sqsubseteq_f B'$, then $A + B \sqsubseteq A' +
B'$. In other words, if systems $A$ and $B$ are individually simulated by $A'$ and
$B'$ on identical alteration of the traces, then the composed
system $A+B$ will be sound with respect to $A'+B'$.
\end{theorem}

\begin{figure}
\centering
\begin{boxedminipage}[c]{.48\textwidth}
%\centering
\inference
[Halt]
{\theta(i) = (s, \pc) & \dec(s,\pc) = H}
{\io{SC}{\theta,m}{H}{}}

\inference
[NonMem]
{\theta(i) = (s, \pc) & \dec(s,\pc) = (\nm, x) \\ \exec(s, \pc, (\nm, x)) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i \coloneqq (s', \pc')], m}{}}

\inference
[Load]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\ld, x, a) \\ \exec(s, \pc, (\ld, x, m(a))) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m}{}}

\inference
[Store]
{\theta(i) = (s,\pc) & \dec(s,\pc) = (\st, a, v) \\ \exec(s, \pc, (\st)) = (\pc', s')}
{\io{SC}{\theta,m}{\theta[i\coloneqq(s',\pc')],m[a\coloneqq v]}{}}
\end{boxedminipage}
\caption{LTS for sequential consistency with $n$ simple processors}
\label{Ref}
\end{figure}

\section{Decomposing a multi-processor system}
\label{sec:store-atomicity}

\begin{figure}
\centering
\includegraphics[scale=.43]{zoom}
\caption{Components of a multi-processors system}
\label{zoom}
\end{figure}

Any multi-processor component can be divided into 3 components as shown in
Figure \ref{zoom}. 
We show the top-level system design at top right, and we zoom in on boxes standing for the
memory system and a single processor ($P_i$).
The processor component $P_i$ can be implemented in a
variety of ways, from a simple in-order issue, to a complex out-of-order issue
core, with aggressive speculation. The memory component is normally implemented
using a hierarchy of caches, in order to increase the performance of the
overall system, since the latency of accessing a memory directly is very slow
compared to accessing a much smaller cache. The component between each
processor and the global memory sub-system contains local buffers, $B_i$, specific to
the processor $P_i$ they are connected to.

\begin{figure}
\centering
\begin{boxedminipage}[c]{.48\textwidth}
\inference
[Load]
{}
{\io{$M_m$}{m}{m}{(\ld\req,a),(\ld\resp,m(a))}}

\inference
[Store]
{}
{\io{$M_m$}{m}{m[a\coloneqq v]}{(\st, a, v)}}

\end{boxedminipage}
\caption{LTS for a simple memory}
\label{$M_m$}
\end{figure}

Independent of the memory models, the memory component always implements the
\emph{store atomicity} property. This property can be understood using the
behavior defined by the operational semantics of Figure \ref{$M_m$}. The
different behaviors permitted by different memory models arise merely because
of the (1) the reordering in the issue of memory instructions by the core $P_i$
if it can issue instructions out-of-order and (2) the properties of the local
buffers $B_i$ connected to each processor $P_i$. For instance, if $B_i$ was a
store buffer, into which the stores from an in-order processor $P_i$ that are
``committed'' are enqueued, and if any ``subsequent'' load from the same
processor $P_i$ reads the latest value from the store buffer, then one can
establish that such a multi-processor system implements the TSO
\cite{x86tsocacm10} memory model.

The memory component is composed of a \emph{hierarchy of caches}, where each
cache node (labeled like ``L1,'' ``L2,'' etc.) communicates only with its
neighbors in the graph.  Each processor has a dedicated L1 cache, from which we
do our best to satisfy all memory requests, to avoid the latency of a
round-trip with main memory.  With an L1 cache miss, we may still find a hit in
a parent cache below the level of main memory, realizing a smaller but still
significant speedup.  Therefore it is the responsibility of the hierarchy of
caches, which form the memory sub-component, to implement the store atomicity
property. In fact, as we will prove in Section \ref{sec:cc}, the purpose of the
cache coherence protocol is to establish this invariant for the memory
sub-component.  Concretely, we have verified a \emph{directory-based} protocol
for coordinating an arbitrary tree of caches, where each node stores a
conservative approximation of its children's states.

As an instance of the above decomposition, we will prove that a multi-processor
system with no local buffering in between the processor and the memory
components indeed implements sequential consistency. We implement a highly
speculative processor which executes instructions and issues loads out of
order, but commits instructions (once some ``verification'' is done) in order.

The processor itself can be decomposed into several components. In the zoomed
in version of Figure \ref{zoom}, we show a highly speculative, out-of-order
issue processor. We have the normal architectural state, such as values of
registers.  Our proofs are generic over \emph{a family of instruction set
architectures}, with parameters for opcode sets and functions for executing
opcodes and decoding them from memory.  Other key components are a \emph{branch
predictor}, which guesses at the control-flow path that a processor will
follow, to facilitate speculation; and a \emph{reorder buffer (ROB)}, which
decides which instructions along that path to try executing ahead of schedule.
Our proofs apply to an arbitrary branch predictor, and they work for any
reorder buffer satisfying a simple semantic condition.

Figure~\ref{proofs} gives the overall proof structure that we employ to verify
the system of Figure~\ref{zoom}. As we will explain later, $P_\text{so}$
represents the speculative out-of-order processor, $M_c$ the cache memory,
$M_m$ the simple memory, $P_\text{ref}$ a simple decoupled processor, and
SC the simple composition of P$^n_\text{ref}+ M_m$ executing each
instruction atomically thus implementing sequential consistency. Notation
$A^n$ is for $n$ copies of system $A$ running in parallel.
$\sqsubseteq$ is a refinement operator, capturing a suitable notion of when
one system implements another.  We go into more detail on the symbols from
the figure in the rest of the paper.

Our current framework establishes theorems of the form ``if system $A$ has a run
with some particular observable behavior, then system $B$ also has a run with
the same behavior.''  In this sense, we say that $A$ correctly implements $B$.
Other important properties, such as \emph{deadlock freedom} for $A$ (which
might get stuck without producing any useful behavior), we leave for future
work.

\begin{figure}
\includegraphics[scale=.45]{proofs}
\caption{Overall proof structure}
\label{proofs}
\end{figure}

\section{Specifying Sequential Consistency}\label{sec:sc}

Our final theorem in this paper establishes that a particular complex hardware
system implements sequential consistency (SC) properly.  We state the theorem
in terms of the trace refinement relation $\sqsubseteq$ developed in the prior
section.  Therefore, we need to define an LTS embodying SC.  The simpler this
system, the better.  We do not need to worry about its performance properties,
since we will prove that an optimized system remains faithful to it.

\begin{figure}
\centering
\begin{boxedminipage}[c]{.48\textwidth}
\inference
[Halt]
{\dec(s,\pc) = H}
{\io{P$_\text{ref}$}{s,\pc}{H}{}}

\inference
[NM]
{\dec(s,\pc) = (\nm, x) & \exec(s, \pc, (\nm, x)) = (\pc', s')}
{\ioe{P$_\text{ref}$}{s,\pc}{s', \pc'}}

\inference
[Ld]
{\dec(s,\pc) = (\ld, x, a) & \exec(s, \pc, (\ld, x, v)) = (\pc', s')}
{\io{P$_\text{ref}$}{s, \pc}
{s', \pc'}{(\ld\req, a), (\ld\resp, v)}}

\inference
[St]
{\dec(s,\pc) = (\st, a, v) & \exec(s, \pc, (\st)) = (\pc', s')}
{\ior{P$_\text{ref}$}{s, \pc}
{s', \pc'}{\st, a, v}}
\end{boxedminipage}

\caption{LTS for a single reference processor (P$_{\text{ref}}$)}
\label{Pref$}
\end{figure}

Figure \ref{Pref$} shows the LTS for a single processor which will be used as a
reference processor.  The definition is parametrized over details of an
\emph{instruction set architecture} (ISA).  In particular, the ISA gives us
some domains of architectural states $s$ (e.g., register files) and of program
counters $\pc$.  A function $\dec(s, \pc)$ figures out which instruction $\pc$
references in the current state, returning the instruction's ``decoded'' form.
A companion function $\exec(s, \pc, \mathit{dec})$ actually executes the
instruction, returning the next program counter $\pc'$ and the new state $s'$.

The legal instruction forms, which are outputs of $\dec$, are $(\nm, x)$, for
an operation not accessing memory; $(\ld, x, a)$, for a memory load from
address $a$; $(\st, a, v)$, for a memory store of value $v$ to address $a$; and
$H$, for a ``halt'' instruction that moves the LTS to state $H$. The parameter
$x$ above represents the rest of the instruction, including the
opcode, registers, constants, \etc{}

The legal inputs to $\exec$ encode both a decoded instruction and any relevant
responses from the memory system.  These inputs are $(\nm, x)$ and $\st$, which
need no extra input from the memory; and $(\ld, x, v)$, where $v$ gives the
contents of the requested memory cell.

We can define sequential consistency using the simple memory defined in Figure
\ref{$M_m$} as follows:
\begin{defn}
$SC \triangleq P_\text{ref} + M_m$
\label{sc}
\end{defn}

We define the initial state of SC as $(\theta_0, m_0)$, where $m_0$ is
some initial memory fixed throughout our development, mapping every
address to value $v_0$; and $\theta_0$ maps every processor ID to
$(s_0, \pc_0)$, using architecture-specific default values $s_0$ and
$\pc_0$.

This LTS encodes Lamport's notion of SC, where processors take turns executing
nondeterministically in a simple interleaving.  We need know very few details
of the ISA to define the SC model.  Our final, optimized system is parametrized
over an ISA in the same way.  In the course of the rest of this paper, we will
define an optimized system $O$ and prove $O \sqsubseteq \text{SC}$.

For a system of $n$ processors, our decoupled SC implementation is
$\text{P$^n_{\text{ref}}$} + M_m$.

\section{Speculative Out-Of-Order Processor}\label{sec:ooo}

The previous section divided a multiprocessor system into two rather
simple components, processors and memories.  We are now ready to begin
developing more complex, optimized versions of each piece, starting
with processors.

We implement a \emph{speculative} processor, which may create many
simultaneous outstanding requests to the memory, as an optimization to
increase parallelism.  Our processor proof is in some sense generic
over correct speculation strategies.  We parametrize over two key
components of a processor design: a \emph{branch predictor}, which
makes guesses about processor-local control flow in advance of
%receiving enough memory responses to resolve
resolving conditional jumps; and a
\emph{reorder buffer}, which decides what speculative instructions (like memory loads)
are worth issuing at which moments (in effect \emph{reordering} later
instructions to happen before earlier instructions have finished).

The branch predictor is the simpler of the two components to model.
We indicate branch-predictor state with metavariable $\ppc$.
The operations on such state are $\curPpc(\ppc)$, to extract the
current program-counter prediction; $\nextPpc(\ppc)$, to ask the
predictor to step forward by one instruction; and $\setBp(\ppc, \pc)$,
to reset prediction to begin at a known-accurate position $\pc$.  It
turns out that we need to impose no explicit correctness criterion on
branch predictors.  The processor uses predictions only as hints, and
it always resets the predictor using $\setBp$ after detecting an
inaccurate hint.

The interface and formal contract of a reorder buffer are more
involved.  We write $\rob$ as a metavariable for reorder buffer
state.  The associated operations are:
\begin{itemize}
\item $\phi$, the state of an empty buffer.
\item $\add(\pc, \rob)$, which appends the program instruction at location $\pc$ to the list of instructions that the buffer is allowed to consider executing.
\item $\get\spc\ld(\rob)$, which returns an optional speculative load to issue
\item $\compute(\rob)$, which models a
step of computation inside the buffer. For instance, it invokes the $\dec$
and $\exec$ functions (as defined for SC) internally to obtain the next program
counter, state, \etc{} (only the speculative states are updated, not the actual states).
\item $\upd\ld(\rob,t,v)$, which informs the buffer that the memory
has returned result value $v$ for the speculative load with tag $t$.
\item $\commit(\rob)$, which returns the next instruction in serial
program order, if we have accumulated enough memory responses to execute it
accurately, or returns
$\epsilon$ otherwise.  When $\commit$ returns an instruction, it also
returns the associated program counter plus the next program counter
we would advance to afterward.
 Furthermore, the instruction is
extended with any relevant response from memory (used only for load
instructions, obtained through $\upd\ld$) and with the new architectural state (\eg{} register
file) after execution.
\item $\retire(\rob)$, which informs the buffer that its $\commit$
instruction was executed successfully, so it is time to move on to the
next instruction.
\end{itemize}

We postpone characterization of correct reorder buffers until after we
present the LTS for speculating processors.

\begin{figure*}[t]
\centering
\begin{boxedminipage}[c]{.95\textwidth}
\inference[Fetch]
{}
{\ioe{P$_\text{so}$}{s,\pc,\rob,\ppc}{s,\pc,\add(\curPpc(\ppc),\rob),\nextPpc(\ppc)}}

\inference[Compute]
{\get\spc\ld(\rob) = \epsilon}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\compute(\rob),\ppc}}

\inference[SpecLd]
{\get\spc\ld(\rob) = (\spc\ld,t,a)}
{\io{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\upd\ld(\compute(\rob),t,v),\ppc}
{(\ld\req, a), (\ld\resp, v)}}

\inference[Abort]
{\commit(\rob) = (\pc',\_,\_) & \pc' \neq \pc}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s,\pc,\phi,\setBp(\ppc,\pc)}}

\inference[Halt]
{\commit(\rob) = H}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{H}}

\inference[Nm]
{\commit(\rob) = (\pc,\pc',(\nm,s'))}
{\ioe{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s',\pc',\retire(\rob),\ppc}}

\inference[St]
{\commit(\rob) = (\pc,\pc',(\st,a,v,s'))}
{\io{\text{P$_\text{so}$}}{s',\pc',\retire(\rob),\ppc}{s,\pc,\rob,\ppc}
{\st, a, v}}

\inference[LdGood]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v,s'))}
{\io{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s',\pc',\retire(\rob),\ppc}{(\ld\req, a), (\ld\resp, v)}}

\inference[LdBad]
{\commit(\rob) = (\pc,\pc',(\ld,x,a,v',s')) & v' \neq v & \exec(s,\pc,(\ld,x,v)) = (\pc'', s'')}
{\io{\text{P$_\text{so}$}}{s,\pc,\rob,\ppc}{s'',\pc'',\phi,\setBp(\ppc,\pc)}{(\ld\req, v), (\ld\resp,v)}}
\end{boxedminipage}

\caption{Speculating, out-of-order issue processor}
\label{Ooo}
\end{figure*}

Figure~\ref{Ooo} defines that LTS,
P$_{\texttt{so}}$. This processor may issue arbitrary speculative
loads, but it only ever \emph{commits} the next instruction in serial
program order. 
We will see the processor issue two kinds of loads, a speculative
load ($\spc\ld$) and a commit or real load ($\ld$).  To maintain SC, every speculative
load must have a matching verification load later on, and we maintain
the illusion that the program only depends on the results of
verification loads, which, along with stores, \emph{must be issued in
serial program order}.

When committing a
previously issued speculative load instruction, the associated speculative
memory load response is verified against the new commit load response. If the
resulting values do not match, the processor terminates all past uncommitted
speculation, by emptying the reorder buffer and resetting the
next predicted program counter in the branch predictor to the correct next value. This step mimics
the mechanism implemented in most modern-day processors, operating on the
conservative assumption that the processor may have performed arbitrary
incorrect speculation after receiving the inaccurate response from
memory.  (Note that here ``inaccurate'' does not mean ``wrong at
the time the response was received,'' but rather ``wrong at commit
time'' or ``correct before, but not anymore.'')

A full processor state is $(s, \pc, \rob, \ppc)$, comprising
architectural state, the program counter, and the reorder buffer and branch
predictor states. Its initial state is given by $(s_0, \pc_0, \phi, \ppc_0)$.
The interface of this processor with memory is identical
to that of the reference processor, but now we can exercise a memory
system's transitions even for speculative loads.

\begin{figure*}[t]
\begin{inv}
If $P_\text{so}$ reaches a state $(s, \pc, \rob, \ppc)$, \ie{}

$\io{P$_\text{so}^{*}$}{s_0,\pc_0,\phi,\ppc_0}{s,\pc,\rob,\ppc}{\eta} \Rightarrow$
\begin{displaymath}
\left\{
\begin{array}{ll}
\commit(\rob) = (\pc, \pc', (\nm, s')) &\Rightarrow
\exists x. \; \dec(s,\pc) = (\nm, x) \wedge \exec(s,\pc, (\nm,x)) =
(\pc', s')\\
\commit(\rob) = (\pc, \pc', (\ld, x, a, v, s')) & \Rightarrow
\dec(s,\pc) = (\ld, x,a) \wedge \exec(s,\pc, (\ld,x,v)) = (\pc', s')\\
\commit(\rob) = (\pc, \pc', (\st, a, v, s')) &\Rightarrow
\dec(s,\pc) = (\st, a,v) \wedge \exec(s,\pc, (\st)) =
(\pc', s')\\
\commit(\rob) = H & \Rightarrow
\dec(s, \pc) = H\\
\end{array}
\right.\end{displaymath}
\label{rob}
\end{inv}
\vspace{-.5cm}
\caption{Correctness of reorder buffer}
\label{robfig}
\end{figure*}

Finally, we impose a general correctness condition on reorder
buffers.  They must maintain Invariant~\ref{rob} in Figure~\ref{robfig}.
Intuitively, whenever the buffer claims in a $\commit$
output that a particular instruction is next to execute, causing
certain state changes, that instruction must really be next in line according
to how the program runs in the SC system, and running it must really cause
those state changes.

When this condition holds, we may conclude the correctness theorem for
out-of-order processors.  We use a trace-transformation function
$\noSpec$ that drops all speculative-load requests and responses.
See Definition~\ref{refines} for a review of how we use such
functions in framing trace refinement.  Intuitively, we prove that any
behavior by the speculating processor can be matched by the simple
processor, with speculative messages erased.

\begin{theorem}
\label{ocorrect}
$\text{P$_\text{so}$} \sqsubseteq_\noSpec$ P$_\text{ref}$
\end{theorem}
\begin{proof}
By induction on P$_\text{so}$ traces, using a proper abstraction
function (which drops the speculative messages and the $\rob$ and $\ppc$ states) to relate the two systems.  Invariant~\ref{rob} is crucial to
relate the reorder buffer's behavior with the simple in-order
execution of P$_\text{ref}$.
\end{proof}

\begin{corollary}
\label{ges}
$\text{P$^n_\text{so}$} \sqsubseteq_{\noSpec^n}$ P$_\text{ref}^n$
\end{corollary}
\begin{proof}
Direct consequence of Theorems~\ref{ocorrect} and~\ref{liftn} (where the
latter is one of our preliminary results about $n$-repetition).
\end{proof}

\section{Cache-Based Memory System}\label{sec:cc}

We now turn our attention toward a more efficient implementation of
memory.  It is physically infeasible to connect many processors
directly to a single memory with low enough latency to handle most
memory requests quickly enough.  Instead, each processor will have its
own \emph{L1 cache} of recently used memory cells.  Whenever possible, we
want to service memory requests from processors' own caches, only
contacting memory for addresses not found in the caches.  To further
avoid contacting main memory, we create additional higher-level
caches, connecting each L1 cache to an L2 cache, potentially
connecting L2 caches to L3, etc., bottoming out in main memory at the
root of our tree. The zoomed-in portion of the memory in Figure~\ref{zoom}
diagrams the basic layout.

Now we have concurrent interaction of many processors with many caches
with main memory, and the relationship with the original $M_m$ system
is far from direct.  However, this intricate concurrent execution is
crucial to hiding the latency of main-memory access.
Figure~\ref{cache} formalizes as an LTS $M_c$ the algorithm we implemented,
for providing the memory abstraction on top of a cache hierarchy.
We have what is called an \emph{invalidating directory-based hierarchical
cache-coherence protocol}.  It is a transliteration of a realistic
algorithm implemented in the Bluespec hardware description
language~\cite{Bluespec:TFRG}.

We now try to explain aspects of this transition system at an
intuitive level.  Thanks to the overall modular proof structure in
this paper, it is not essential to understand the details of the cache
system, to understand any other part of our development.  Therefore,
some readers may want to skip this explanation upon first reading,
skipping to the next section for a key theorem relating $M_c$ and $M_m$.

We describe a state of the system using 8 fields: $\dt$, $\ch$, $\s$, $\dst$,
$\wt$, $\dwt$ which are summarized in Figure~\ref{fields}.
We use $\parent(c,p)$ to denote that $p$ is the parent of $c$ and $\leaf(c)$ to 
denote that $c$ is a leaf or L1 cache.

\begin{figure}
\centering
\begin{tabular}{|l|p{5.5cm}|}
\hline
$\dt(c,a)$ & Data in cache $c$ for address $a$\\
$\s(c,a)$ & Coherence state of cache $c$ for address $a$\\
$\dst(p,c,a)$ & Directory state of cache $p$ for child $c$ and address $a$\\
$\wt(c,a)$ & Gives cache $c$'s desired coherence state to upgrade to (for address $a$), or $\epsilon$ if none\\
$\dwt(p,c,a)$ & Gives cache $p$'s desired coherence state for its child $c$ to downgrade to (for address $a$), or $\epsilon$ if none\\
$\ch$ & Channels between nodes in the system\\
\hline
\end{tabular}
\caption{Fields in a cache-based memory system}
\label{fields}
\end{figure}

Fields $\dt, \s$ are maps over pairs of cache names and memory addresses.
The $\dt$ map assigns each pair a memory-cell value, and the $\s$ map
assigns each pair a coherence state.  A coherence state is $M, S$ or
$I$, broadly representing permissions to modify, read, or do nothing
with an address, respectively, the decreasing permissions denoted by $M > S > I$. This interpretation of permissions is
accurate at the leaf caches connected to the processors (\ie{} L1 caches),
but the interpretation is subtly different at the higher level caches.

Mapping $\wt$ goes from cache name and address to either $\epsilon$ or a
coherence state.  It denotes what permission the cache is waiting for,
if any.  That is, a cache may have decided to \emph{upgrade} its
coherence state for some address to a more permissive value, but the
cache is waiting for acknowledgment from another cache before
finalizing the upgrade.

Since we have a directory-based protocol, the $\s$ and
$\wt$ maps have mirror-image counterparts $\dst$ and $\dwt$.
The former maps triples of parent name, child name, and address to
coherence state. It represents the parent's notion of the
coherence state of the child for that address. We later prove that this notion
is always conservative, \ie{} if the parent assumes that a child does not have
a particular permission, then it is guaranteed in this system that the child
will not have that permission.  The $\dwt$ field also represents a mapping from parent
name, child name, and address to either $\epsilon$ or coherence state. It
denotes whether the parent is waiting for any downgrade response from the
corresponding child, and if so, the coherence state that the child must
downgrade to as well.

All communication channels in the system are represented by field $\ch$.
The channels are asymmetric as
follows. Channels going from parent to children (indexed by the names of the
parent and child) maintain FIFO order of messages between the same parent-child
pair. They carry both request messages (asking the child to downgrade) and
response messages (acknowledging that the child can upgrade). Channels going
from children to their parents are further split into request and response
channels, thus indexed by tuples (child name, parent name, request or response).
These channels need not maintain the FIFO order for messages between the same
child-parent pair (notice the use of $\coloncolon$ to add and remove elements
from the parent to child channel, as opposed to $\uplus$ for the 
child to parent channel). The reason is that only one request or one response is
sent for a particular address between a child-parent pair -- this invariant can
be proved for the LTS in  Figure \ref{cache}. The order between a request and
response message sent from the same child to its parent is lost, but our
protocol ensures that nothing bad happens as a result.  Notice rules
ParentRecvRsp and ParentRecvReq in Figure \ref{cache}, each of which looks at the
previous coherence state of the incoming message and compares it with its own
directory state to decide whether to process the message.

%\medskip

Here is an intuition on how the transitions work in the common case.  A cache
can decide, spontaneously, to upgrade its coherence state, in which case it
sends an upgrade request to its parent. The parent then makes a local decision
on whether to send a response to the requesting child or not, based on its
directory approximation and its own coherence state $\s$. If $\s$ is lower than
the requested upgrade, then it cannot handle the request, and instead must
decide to upgrade $\s$.  Once the parent's $\s$ is not lower than the requested
upgrade, it makes sure that the rest of its children are ``compatible'' with
the requested upgrade (given by $\compat$ definition below).  If not, the
parent must send requests to the incompatible children to downgrade. Finally, when
the $\s$'s upgrade and children's downgrade responses are all received, the
original request can be responded to. The transitions for sending requests are
allowed to happen spontaneously without any trigger. An L1 cache can process a
load request or a store request only if it is in at least the \Sh{} state or
\Mo{} state, respectively, for the requested address; otherwise it must
spontaneously request to upgrade.

\begin{defn}
\begin{displaymath}
\compat(p, c, x, a) = \left\{
\begin{array}{ll}
\forall c' \neq c. \; \dst(p,c',a) = I &: x = M\\
\forall c' \neq c. \; \dst(p,c',a) \le S &: x = S\\
\end{array}
\right.
\end{displaymath}
\end{defn}

The reason for having separate request and response channels for the same
child-parent pair is to avoid deadlocks. Throughout the protocol, we maintain
the invariant
that a downgrade request from the parent always gets priority, \ie{} if a child
is waiting for an upgrade response, it must still try to satisfy a downgrade
request from the parent. If both the parent and the child are waiting
for their respective responses, there will be a deadlock. This tie cannot
be broken arbitrarily: if the parent concedes and starts processing a child's
request, then it may have to send downgrade requests to other children to make
the states compatible, but the other children may have sent upgrade requests
and remain waiting, leading to a deadlock. The only way to break the
vicious waiting cycle is for the children to concede and respond when they get
downgrade requests. Of course, they have to downgrade the states of their own
children first, and this process goes on recursively until the downgrade requests
reach the leaf caches, which can respond immediately.

A complication arises because a cache can voluntarily decide to downgrade its
state.  This transition is used to model invalidation of cache lines to make
room for a different location (we permit a cache to downgrade from $M$ to $S$
instead of only to $I$).
As a result, the parent's $\dst$ and the
corresponding $\s$ of the child may go out of sync, leading to the parent
requesting a child to downgrade when it already has. To handle this situation,
the child has to drop the downgrade request when it has already downgraded to
the required state (Rule DropReq in Figure \ref{cache}), to avoid deadlocks
by not dequeuing the request.

%Dropping the request
%removes the entry from the buffer, allowing the child to process further messages
%from the parent -- otherwise there would have been a deadlock.

\begin{figure*}
\centering

\begin{boxedminipage}[c]{.95\textwidth}

\textbf{Processor/Memory Interface}

\inference[Load]
{\s(c,a) \geq S}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt, \ch, \s, \dst, \wt, \dwt}{c,((\ld\req, a),(\ld\resp,v))}}
\vspace{.1in}
\inference[Store]
{s(c,a) \geq M}
{\io{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt[(c, a) \coloneqq v], \ch, \s, \dst, \wt, \dwt}{c,(\st,a,v)}}

\textbf{Child Upgrade}

\inference[ChildSendReq]
{\parent(c,p) & \s(c,a) < x & \wt(c,a) = \epsilon}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst, \wt, \\ \dwt}
{\dt, \ch[ (c, p, \req) \coloneqq (a, \s(c, a), x) \uplus \ch(c,p,\req)], \s, \dst,\\
\wt[(c,a)\coloneqq x], \dwt}}
\vspace{.1in}
\inference[ParentRecvReq]
{\parent(c,p) & \ch(c, p, \req) = \{(a, y, x)\} \uplus \rs & \s(p, a) \geq x
\\ \compat(p, c, x, a) & \dwt(p,c,a) = \epsilon & \dst(p,c,a) \le y}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst,\\ \wt, \dwt}
{\dt, \ch[(c,p,\req) \coloneqq \rs][(p, c) \coloneqq (\resp, (a, \dst(p,c,a), x,\\ \;\;\;\;
\ite{\dst(p,c,a) = I}{\dt(p,a)}{\_})) \coloncolon \ch(p,c)], \\ \s, \dst[(p,c, a)\coloneqq x],
\wt, \dwt}}
\vspace{.1in}
\inference[ChildRecvRsp]
{\parent(c,p) & \ch(p, c) = \rs \coloncolon (\resp, (a, y, x, v))}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst, \wt, \\ \dwt}
{\dt[(c,a) \coloneqq \ite{y = I}{v}{\dt(c,a)}], \ch[(p,c) \coloneqq \rs], \s[(c,a) \coloneqq x], \dst,\\
\wt[(c,a)\coloneqq \ite{\wt(c,a) \le x}{\epsilon}{\wt(c,a)}], \dwt}}

\textbf{Parent Downgrade}

\inference[ParentSendReq]
{\parent(c,p) & \dst(p,c,a) > x & \dwt(p,c,a) = \epsilon}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst, \wt, \\ \dwt}
{\dt, \ch[(p, c)\coloneqq (\req, (a, \dst(p, c, a), x)) \coloncolon \ch(p,c)], \s, \dst,\\
\wt, \dwt[(p, c,a)\coloneqq x]}}
\vspace{.1in}
\inference[ChildRecvReq]
{\parent(c,p) & \ch(p, c) = \rs \coloncolon (\req, (a, y, x)) & (\forall i. \; \parent(i, c) \Rightarrow \dst(c, i, a) \le x)
& \s(c,a) > x}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst,\\ \wt, \dwt}
{\dt, \ch[(p,c) \coloneqq \rs][(c, p,\resp) \coloneqq (a, \s(c,a), x,\\ \;\;\;\;
\ite{\dst(c,a) = M}{\dt(c,a)}{\_}) \uplus \ch(c,p,\resp)], \\ \s[(c,a)\coloneqq x], \dst,
\wt, \dwt}}
\vspace{.1in}
\inference[ParentRecvRsp]
{\parent(c,p) & \ch(c,p,\resp) = \{(a, y, x, v)\} \uplus \rs & \dst(p,c,a) = y}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst, \wt, \\ \dwt}
{\dt[(p,a) \coloneqq \ite{y = M}{v}{\dt(p,a)}], \ch[(c,p,\resp) \coloneqq \rs], \s, \dst[(p,c,a) \coloneqq x],\\
\wt, \dwt[(p,c,a)\coloneqq \ite{\dwt(p,c,a) \ge x}{\epsilon}{\dwt(p,c,a)}]}}

\textbf{Voluntary downgrade for replacement}

\inference[VolResp]
{\parent(c,p) & (\forall i. \; \parent(i, c) \Rightarrow \dst(c, i, a) \le x)
& \s(c,a) > x}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst,\\ \wt, \dwt}
{\dt, \ch[(c, p,\resp) \coloneqq (a, \s(c,a), x,\\ \;\;\;\;
\ite{\s(c,a) = M}{\dt(c,a)}{\_}) \uplus \ch(c,p,\resp)], \\ \s[(c,a)\coloneqq x], \dst,
\wt, \dwt}}

\textbf{Dropping request because of voluntary downgrade}

\inference[DropReq]
{\parent(c,p) & \ch(p, c) = \rs \coloncolon (\req, (a, y, x)) & 
 \s(c,a) \le x}
{\ioxe{$M_c$}
{\dt, \ch, \s, \dst, \wt, \dwt}
{\dt, \ch[(p,c) \coloneqq \rs], \s, \dst,
\wt, \dwt}}
\end{boxedminipage}

\caption{LTS for cache-coherent shared-memory system}
\label{cache}
\end{figure*}

\section{Proof that the MSI Protocol Preserves Store Atomicity}\label{sec:ccproof}
\label{safety}

We must prove the following theorem, \ie{} the cache-based system is sound with
respect to the simple memory.
\begin{theorem}
\label{ccorrect}
$M_c \sqsubseteq M_m$
\end{theorem}

The key property we need for this proof is what we define as
``store atomicity'' below. Throughout this section, we say
\emph{time} to denote the number of transitions that occurred before reaching
the specified state.

\begin{defn}
Store atomicity is defined as the following property:
For a load (or speculative load) request $\toM(\ld, a)$ (or $\toM(\spc\ld, t, a)$)
from any processor, the response sent at time $T$
$\toP(\ld, v)$ (or $\toP(\spc\ld, t, v)$) should be such that
\begin{enumerate}
\item $v = v_0$ (the initial value of any memory address) and no store
  request $\toM(\st, a, v')$ has been processed at any time $T'$ such
  that $T' < T$ or
\item There is a store request $\toM(\st, a, v)$ that was processed at time $T_q$ such that
$T_q < T$ and no other store
request $\toM(\st, a, v')$ was processed at any time $T'$ such that $T_q < T' < T$.
\end{enumerate}
\label{sa}
\end{defn}

We can easily show that any system obeying store atomicity
is sound with respect to $M_m$. Intuitively, the store atomicity
property precisely describes the property of a memory, \viz{} that if a store
happens, then that location gets updated such that any subsequent load gets
back the updated value.

The proof that $M_c$ obeys store atomicity is involved enough to deserve a
section's worth of explanation, which we now turn to.

We first reduce the proof of store atomicity to several lemmas. We then
present a key insight that we discovered for proving these lemmas in the form
of three fundamental invariants. We believe that other invalidation-based
protocols with inclusive caches (\ie{} those cache hierarchies where a parent
cache contains all the locations that any of its children contains) like MESI,
MOSI, and MOESI also satisfy our invariants and lemmas, making their verification
easier by reusing most of our MSI proof; we omit the discussion of extension of
this proof to other protocols for space reasons.

%\adamc{...and maybe we don't, for space reasons. \texttt{:-)}}

%We use the terms write to mean handling a store request, \ie{} executing the
%StoreReq transition and read to mean handling a load request, \ie{} executing
%the LoadReq transition.

The lemmas and invariants presented below hold after any number of transitions
starting from the initial state.

While we have proofs in Coq for each lemma, our discussion will be mainly
informal for brevity. All the proofs can be found in our supplementary material.
%\adamc{or just remind of the source code URL given earlier}

%\subsection{Invariants}
%\label{firstLevel}
Intuitively our protocol is correct because it is guaranteed that any cache that attempts to
handle a load request has the ``latest'' value.

\begin{lemma}
\textit{latestValue}:
At any time $T$, a cache $c$ that is \clean{} (defined below) for an address
$a$ will have the most up-to-date value for that address, \ie{}
\begin{enumerate}
\item $\dt(c,a) = v_0$ and no store request $\toM(\st, a, v)$ has been processed at
any time $T'$ such that $T' < T$ or
\item There is a store request $\toM(\st, a, v)$ that was processed at time $T_q$ such that
$T_q < T \wedge \dt(c,a) = v$ and no other store
request $\toM(\st, a, v')$ was processed at any time $T'$ such that $T_q < T' < T$.
\end{enumerate}
\label{latestValue}
\end{lemma}

\vspace{-.15in}

\begin{defn}
\textit{clean}: A cache is said to be \textit{clean} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for any of its children is at most \Sh.
\label{clean}
\end{defn}

It is pretty intuitive to prove the store atomicity property given
Lemma~\ref{latestValue}. A leaf cache can respond to a request only when it
has the required permission, at which point it is guaranteed to have the latest
value, because it is \textit{clean} (as it has no children).

Lemma~\ref{latestValue} requires further decomposition into more
lemmas.

\begin{lemma}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$ such
that neither is an ancestor of the other in the cache-tree hierarchy (where
the ancestor relation is the reflexive-transitive closure of the parent relation),
for each address $a$, $\s(c_1,a)$ and $\s(c_2,a)$ are
\textit{compatible} (defined below).
\label{nonAncestorCompatible}
\end{lemma}
\begin{defn}
\textit{compatible}: Values of states $x_1$ and $x_2$ are \textit{compatible}
iff whenever either $x_1$ or $x_2$ is \Mo{}, the other is \In.
\label{compatible}
\end{defn}

It is easy to see that any two states $y_1 \le x_1$ and $y_2 \le x_2$ are also
compatible if $x_1$ and $x_2$ are compatible.

\begin{lemma}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall T. \; T_s \le T \le T_r$ where $T_s$ is the time of sending the data and
$T_r$ the time of receiving the data), no cache can process a store request for
$a$, and the data must be sent from a \textit{clean} cache.
\label{noTransitWrite}
\end{lemma}

\begin{lemma}
\textit{conservative}: The state of an address in a cache is never greater than
the corresponding entry in its parent's directory.
\label{conservative}
\end{lemma}

\begin{lemma}
\textit{localCompatibility1}:
The state of a cache is never less than its directory entry for any of
its children.
\label{localCompatibility1}
\end{lemma}
\begin{lemma}
\textit{localCompatibility2}: Let $c_1$ and $c_2$ be distinct caches with the
same parent $p$. Then, $\s(c_1, a)$ and $\s(c_2, a)$ are
\textit{compatible}.
\label{localCompatibility2}
\end{lemma}

Lemmas~\ref{conservative} and~\ref{localCompatibility1} imply a
$\ge$ relation between the states of a parent and its child for any address.
Since the ancestor relation is the reflexive-transitive closure of the parent
relation, the $\ge$ relation is preserved between the states of two caches $p$
and $c$ where $p$ is the ancestor of $c$.

Any two caches $c_1, c_2$ neither of which is an ancestor of the other have a
lowest common ancestor $x$, which has
two distinct children $x_1$ and $x_2$ such that they are the ancestors of $c_1$
and $c_2$ respectively. For any address $a$, $\s(x_1, a)$ and
$\s(x_2, a)$ are compatible by Lemma~\ref{localCompatibility2}. From
the previous paragraph about the relation between the states of a cache and its
ancestor, it follows that $c_1$ and $c_2$ are also compatible, proving
Lemma~\ref{nonAncestorCompatible}.

Using Lemmas~\ref{nonAncestorCompatible} and~\ref{noTransitWrite}, one can
prove Lemma~\ref{latestValue}. The proof is by induction on time $T$; we assume
that any \clean{} (node, address) pair has the latest value until time
$T$. Consider a (node, addr) pair at time $T+1$.
If the pair was \clean{} at $T$, then no cache,
save itself (because of Lemma~\ref{nonAncestorCompatible}), can process any
store request for that address, thus continuing to be \clean{} and having the latest value.
And if the pair was not \clean{} at time $T$, it will receive data from a cache
that is \clean{} for the (node, address) pair, and no store
request can be processed when the data is in transit (by
Lemma~\ref{noTransitWrite}). Therefore, by the induction hypothesis the cache will
continue to have the latest value whenever it is \clean{} for an address at time $T+1$.

\subsection{Key Insight for Proving Lemmas}

We present what we believe to be the key insight for proving any
cache-coherence protocol.  In fact, the request/response message semantics for
the MSI protocol was designed to satisfy the following three invariants.

\begin{inv}
\textit{stateChange}:
\begin{enumerate}
\item A cache downgrades its state for an address iff it sends a response for that address.
\item A parent upgrades the directory entry for an address for one of its
children iff it sends a response for that child and address.
\item The state or directory entry changes only on sending or
receiving a response $(\resp, (a, y, x, d))$, to $x$.
\end{enumerate}
\label{stateChange}
\end{inv}

\vspace{-.1in}

Invariant~\ref{stateChange} ensures that a response message acts as a sync
between a parent's directory entry for a child and the state of the child, as
both the sender and the receiver change their respective directory entry or
state to the same $x$ in a response $(\resp, (a, y, x, d))$

\begin{inv}
\textit{noCross}: A response from a parent to its child for an address cannot
be in transit while another response from the same child to the parent for the
same address is in transit.
\label{noCross}
\end{inv}

\begin{inv}
\textit{respFIFO}: Responses are received in the order in which they are sent,
for a source-destination pair, for the same address.
\label{respFIFO}
\end{inv}

\begin{figure}
\centering
\includegraphics[scale=.45]{resps}
\caption{An example showing a sequence of responses between a child and its parent}
\label{resps}
\end{figure}

Figure~\ref{resps} shows an example of the effect of
Invariants~\ref{stateChange}, \ref{noCross} and \ref{respFIFO}. The
directory entry of the parent or the state of the child changes on sending and
receiving a response, and the new state is shown at the points of sending or
receiving responses.  Both the cache and its parent follow the same sequence of
state and directory entry changes, but the receiver always lags slightly behind
the sender.  The receiver cannot send any response before receiving because of
Invariant~\ref{noCross}, thereby eliminating any scenarios where the lag
complicates reasoning.

Using these three invariants (Invariants~\ref{stateChange}, \ref{noCross} and
\ref{respFIFO}), it is easy to see that Lemma~\ref{conservative} holds.
Intuitively, going back to Figure~\ref{resps}, we can see that at all times,
either there is no response in flight, only responses from the parent are in
flight, or only responses from the child are in flight. A response from child
downgrades the state, and a response from parent upgrades the directory. Since
the other side simply lags behind, Lemma~\ref{conservative} holds.
Since Invariants~\ref{stateChange}, \ref{noCross} and \ref{respFIFO} deal with
just one cache's state for a single address and the corresponding directory
entry, these properties can also potentially be verified using model-checking.
Thus, if some other protocol is verified to obey Invariants~\ref{stateChange},
\ref{noCross} and \ref{respFIFO} using model-checking, then
Lemma~\ref{conservative} is already automatically verified for that
protocol.

Analyzing a parent-child pair will suffice in proving
Lemmas~\ref{localCompatibility1} and \ref{localCompatibility2}.  Proof of
Lemma~\ref{noTransitWrite} is similar to the proof of
Lemma~\ref{nonAncestorCompatible}.

%Figure \ref{something} gives the structure of the proof for Theorem \ref{ccorrect}
%\begin{figure*}
%\begin{prooftree}
%\AxiomC{stateChange}
%\AxiomC{noCross}
%\AxiomC{respFifo}
%\TrinaryInfC{conservative}
%\AxiomC{localCompatibility1}
%\AxiomC{localCompatibility2}
%\TrinaryInfC{nonAncestorCompatible}
%\AxiomC{noTransitWrite}
%\BinaryInfC{latestValue}
%\AxiomC{$\ldots$ (trivial)}
%\BinaryInfC{StoreAtomicity}
%\UnaryInfC{$M_c \sqsubseteq M_m$}
%\end{prooftree}
%\caption{Proving $M_c \sqsubseteq M_m$}
%\label{something}
%\end{figure*}

\section{The Final Result}\label{sec:finalresult}

With our two main results about optimized processors and memories, it
is now straightforward to complete the correctness proof of the
composed optimized system.

First, we need to know that, whenever the simple memory can generate
some trace of messages, it could also generate the same trace with all
speculative messages removed.  We need this property to justify the
introduction of speculation, during our final series of refinements
from the optimized system to SC.

\begin{theorem}
\label{mcorrect}
$M_m \sqsubseteq_{\noSpec^n} M_m$
\end{theorem}
\begin{proof}
By induction on traces, with an identity abstraction function.
\end{proof}

That theorem turns out to be the crucial ingredient to justify placing
a speculative processor in-context with simple memory.

\begin{theorem}
\label{ocorrect1}
$\text{P$^n_\text{so}$} + M_m \sqsubseteq \text{P$^n_\text{ref}$} + M_m$
\end{theorem}
\begin{proof}
Follows from Theorem \ref{liftplus} (one of our preliminary results
about $+$), Corollary \ref{ges}, and Theorem \ref{mcorrect}.
\end{proof}

The last theorem kept the memory the same while refining the
processor.  The next one does the opposite, switching out memory.

\begin{theorem}
\label{ocorrect2}
$\text{P$^n_\text{so}$} + M_c \sqsubseteq \text{P$^n_\text{so}$} + M_m$
\end{theorem}
\begin{proof}
Follows from Theorems \ref{ccorrect} and \ref{liftplus} plus reflexivity of $\sqsubseteq$ (Theorem \ref{transitive}).
\end{proof}

\begin{theorem}
\label{ofull}
$\text{P$^n_\text{so}$} + M_c \sqsubseteq SC$
\end{theorem}
\begin{proof}
We twice apply $\sqsubseteq$ transitivity (Theorem \ref{transitive}) to
connect Theorems \ref{ocorrect2} and \ref{ocorrect1}, and use Definition \ref{sc} of $SC$
\end{proof}