### 2014-10-30 TransitionsUsefulThird.tex

\section{Cache-Based Memory System}\label{sec:cc}

We now turn our attention toward a more efficient implementation of
memory.  It is physically infeasible to connect many processors
directly to a single memory with low enough latency to handle most
memory requests quickly enough.  Instead, each processor will have its
own \emph{L1 cache} of recently used memory cells.  Whenever possible, we
want to service memory requests from processors' own caches, only
contacting memory for addresses not found in the caches.  To further
avoid contacting main memory, we create additional higher-level
caches, connecting each L1 cache to an L2 cache, potentially
connecting L2 caches to L3, etc., bottoming out in main memory at the
root of our tree. The zoomed-in portion of the memory in Figure~\ref{zoom}
diagrams the basic layout.

Now we have concurrent interaction of many processors with many caches
with main memory, and the relationship with the original $M_m$ system
is far from direct.  However, this intricate concurrent execution is
crucial to hiding the latency of main-memory access.
Figure~\ref{cache} formalizes as an LTS $M_c$ the algorithm we implemented,
for providing the memory abstraction on top of a cache hierarchy.
We have what is called an \emph{invalidating directory-based hierarchical
cache-coherence protocol}.  It is a transliteration of a realistic
algorithm implemented in the Bluespec hardware description
language~\cite{Bluespec:TFRG}.

We now try to explain aspects of this transition system at an
intuitive level.  Thanks to the overall modular proof structure in
this paper, it is not essential to understand the details of the cache
system, to understand any other part of our development.  Therefore,
some readers may want to skip this explanation upon first reading,
skipping to the next section for a key theorem relating $M_c$ and $M_m$.

We describe a state of the system using 8 fields: $\dt$, $\ch$, $\s$, $\dst$,
$\wt$, $\dwt$, $\inp$, $\outp$. The $\inp$ and $\outp$ sets are the interfaces
to the processors and are exactly the same as in the simple memory LTS
(Figure~\ref{$M_m$}). The rest are summarized in Figure~\ref{fields}.

%We describe a state of the system using 6 fields: $\dt$, $\ch$, $\s$, $\dst$,
%$\wt$, $\dwt$ which are summarized in Figure~\ref{fields}.
We use $\parent(c,p)$ to denote that $p$ is the parent of $c$ and $\leaf(c)$ to 
denote that $c$ is a leaf or L1 cache.

\begin{figure}
\centering
\begin{tabular}{|l|p{5.5cm}|}
\hline
$\dt(c,a)$ & Data in cache $c$ for address $a$\\
$\s(c,a)$ & Coherence state of cache $c$ for address $a$\\
$\dst(p,c,a)$ & Directory state of cache $p$ for child $c$ and address $a$\\
$\wt(c,a)$ & Gives cache $c$'s desired coherence state to upgrade to (for address $a$), or $\epsilon$ if none\\
$\dwt(p,c,a)$ & Gives cache $p$'s desired coherence state for its child $c$ to downgrade to (for address $a$), or $\epsilon$ if none\\
$\ch$ & Channels between nodes in the system\\
\hline
\end{tabular}
\caption{Fields in a cache-based memory system}
\label{fields}
\end{figure}

Fields $\dt, \s$ are maps over pairs of cache names and memory addresses.
The $\dt$ map assigns each pair a memory-cell value, and the $\s$ map
assigns each pair a coherence state.  A coherence state is $M, S$ or
$I$, broadly representing permissions to modify, read, or do nothing
with an address, respectively, the decreasing permissions denoted by $M > S > I$. This interpretation of permissions is
accurate at the leaf caches connected to the processors (\ie{} L1 caches),
but the interpretation is subtly different at the higher level caches.

Mapping $\wt$ goes from cache name and address to either $\epsilon$ or a
coherence state.  It denotes what permission the cache is waiting for,
if any.  That is, a cache may have decided to \emph{upgrade} its
coherence state for some address to a more permissive value, but the
cache is waiting for acknowledgment from another cache before
finalizing the upgrade.

Since we have a directory-based protocol, the $\s$ and
$\wt$ maps have mirror-image counterparts $\dst$ and $\dwt$.
The former maps triples of parent name, child name, and address to
coherence state. It represents the parent's notion of the
coherence state of the child for that address. We later prove that this notion
is always conservative, \ie{} if the parent assumes that a child does not have
a particular permission, then it is guaranteed in this system that the child
will not have that permission.  The $\dwt$ field also represents a mapping from parent
name, child name, and address to either $\epsilon$ or coherence state. It
denotes whether the parent is waiting for any downgrade response from the
corresponding child, and if so, the coherence state that the child must
downgrade to as well.

All communication channels in the system are represented by field $\ch$.
The channels are asymmetric as
follows. Channels going from parent to children (indexed by the names of the
parent and child) maintain FIFO order of messages between the same parent-child
pair. They carry both request messages (asking the child to downgrade) and
response messages (acknowledging that the child can upgrade). Channels going
from children to their parents are further split into request and response
channels, thus indexed by tuples (child name, parent name, request or response).
These channels need not maintain the FIFO order for messages between the same
child-parent pair (notice the use of $\coloncolon$ to add and remove elements
from the parent to child channel, as opposed to $\uplus$ for the 
child to parent channel). The reason is that only one request or one response is
sent for a particular address between a child-parent pair -- this invariant can
be proved for the LTS in  Figure \ref{cache}. The order between a request and
response message sent from the same child to its parent is lost, but our
protocol ensures that nothing bad happens as a result.  Notice rules
ParentRecvRsp and ParentRecvReq in Figure \ref{cache}, each of which looks at the
previous coherence state of the incoming message and compares it with its own
directory state to decide whether to process the message.

Here is an intuition on how the transitions work in the common case.  A cache
can decide, spontaneously, to upgrade its coherence state, in which case it
sends an upgrade request to its parent. The parent then makes a local decision
on whether to send a response to the requesting child or not, based on its
directory approximation and its own coherence state $\s$. If $\s$ is lower than
the requested upgrade, then it cannot handle the request, and instead must
decide to upgrade $\s$.  Once the parent's $\s$ is not lower than the requested
upgrade, it makes sure that the rest of its children are ``compatible'' with
the requested upgrade (given by $\compat$ definition below).  If not, the
parent must send requests to the incompatible children to downgrade. Finally, when
the $\s$'s upgrade and children's downgrade responses are all received, the
original request can be responded to. The transitions for sending requests are
allowed to happen spontaneously without any trigger. An L1 cache can process a
load request or a store request only if it is in at least the \Sh{} state or
\Mo{} state, respectively, for the requested address; otherwise it must
spontaneously request to upgrade.

\begin{defn}
\begin{displaymath}
\compat(p, c, x, a) = \left\{
\begin{array}{ll}
\forall c' \neq c. \; \dst(p,c',a) = I &: x = M\\
\forall c' \neq c. \; \dst(p,c',a) \le S &: x = S\\
\end{array}
\right.
\end{displaymath}
\end{defn}

The reason for having separate request and response channels for the same
child-parent pair is to avoid deadlocks. Throughout the protocol, we maintain
the invariant
that a downgrade request from the parent always gets priority, \ie{} if a child
is waiting for an upgrade response, it must still try to satisfy a downgrade
request from the parent. If both the parent and the child are waiting
for their respective responses, there will be a deadlock. This tie cannot
be broken arbitrarily: if the parent concedes and starts processing a child's
request, then it may have to send downgrade requests to other children to make
the states compatible, but the other children may have sent upgrade requests
and remain waiting, leading to a deadlock. The only way to break the
vicious waiting cycle is for the children to concede and respond when they get
downgrade requests. Of course, they have to downgrade the states of their own
children first, and this process goes on recursively until the downgrade requests
reach the leaf caches, which can respond immediately.

A complication arises because a cache can voluntarily decide to downgrade its
state.  This transition is used to model invalidation of cache lines to make
room for a different location (we permit a cache to downgrade from $M$ to $S$
instead of only to $I$).
As a result, the parent's $\dst$ and the
corresponding $\s$ of the child may go out of sync, leading to the parent
requesting a child to downgrade when it already has. To handle this situation,
the child has to drop the downgrade request when it has already downgraded to
the required state (Rule DropReq in Figure \ref{cache}), to avoid deadlocks
by not dequeuing the request.

\input{TransitionsUsefulCache1}

\section{Proof that the MSI Protocol Preserves Store Atomicity}\label{sec:ccproof}
\label{safety}

We must prove the following theorem, \ie{} the cache-based system is sound with
respect to the simple memory.
\begin{theorem}
\label{ccorrect}
$M_c \sqsubseteq M_m$
\end{theorem}

The key property we need for this proof is what we define as
``store atomicity'' below. Throughout this section, we say
\emph{time} to denote the number of transitions that occurred before reaching
the specified state.

\begin{defn}
Store atomicity is defined as the following property:
For a load (or speculative load) request $\toM(\ld, a)$ (or $\toM(\spc\ld, t, a)$)
from any processor, the response sent at time $T$
$\toP(\ld, v)$ (or $\toP(\spc\ld, t, v)$) should be such that
\begin{enumerate}
\item $v = v_0$ (the initial value of any memory address) and no store
  request $\toM(\st, a, v')$ has been processed at any time $T'$ such
  that $T' < T$ or
\item There is a store request $\toM(\st, a, v)$ that was processed at time $T_q$ such that
$T_q < T$ and no other store
request $\toM(\st, a, v')$ was processed at any time $T'$ such that $T_q < T' < T$.
\end{enumerate}
\label{sa}
\end{defn}

We can easily show that any system obeying store atomicity
is sound with respect to $M_m$. Intuitively, the store atomicity
property precisely describes the property of a memory, \viz{} that if a store
happens, then that location gets updated such that any subsequent load gets
back the updated value.

The proof that $M_c$ obeys store atomicity is involved enough to deserve a
section's worth of explanation, which we now turn to.

We first reduce the proof of store atomicity to several lemmas.
We believe that other invalidation-based
protocols with inclusive caches (\ie{} those cache hierarchies where a parent
cache contains all the locations that any of its children contains) like MESI,
MOSI, and MOESI also satisfy our lemmas, making their verification
easier by reusing most of our MSI proof; we omit the discussion of extension of
this proof to other protocols for space reasons.

The lemmas
presented below hold after any number of transitions
starting from the initial state.

While we have proofs in Coq for each lemma, our discussion will be mainly
informal for brevity. All the proofs can be found in our supplementary material.

Intuitively our protocol is correct because it is guaranteed that any cache that attempts to
handle a load request has the ``latest'' value.

\begin{lemma}
\textit{latestValue}:
At any time $T$, a cache $c$ that is \clean{} (defined below) for an address
$a$ will have the most up-to-date value for that address, \ie{}
\begin{enumerate}
\item $\dt(c,a) = v_0$ and no store request $\toM(\st, a, v)$ has been processed at
any time $T'$ such that $T' < T$ or
\item There is a store request $\toM(\st, a, v)$ that was processed at time $T_q$ such that
$T_q < T \wedge \dt(c,a) = v$ and no other store
request $\toM(\st, a, v')$ was processed at any time $T'$ such that $T_q < T' < T$.
\end{enumerate}
\label{latestValue}
\end{lemma}

\vspace{-.15in}

\begin{defn}
\textit{clean}: A cache is said to be \textit{clean} for an address $a$ if and only if
the state of the cache for that address is at least \Sh{} and the directory 
state for any of its children is at most \Sh.
\label{clean}
\end{defn}

It is pretty intuitive to prove the store atomicity property given
Lemma~\ref{latestValue}. A leaf cache can respond to a request only when it
has the required permission, at which point it is guaranteed to have the latest
value, because it is \textit{clean} (as it has no children).

Lemma~\ref{latestValue} requires further decomposition into more
lemmas.

\begin{lemma}
\textit{nonAncestorCompatible}: For two distinct caches $c_1$ and $c_2$ such
that neither is an ancestor of the other in the cache-tree hierarchy (where
the ancestor relation is the reflexive-transitive closure of the parent relation),
for each address $a$, $\s(c_1,a)$ and $\s(c_2,a)$ are
\textit{compatible} (defined below).
\label{nonAncestorCompatible}
\end{lemma}
\begin{defn}
\textit{compatible}: Values of states $x_1$ and $x_2$ are \textit{compatible}
iff whenever either $x_1$ or $x_2$ is \Mo{}, the other is \In.
\label{compatible}
\end{defn}

It is easy to see that any two states $y_1 \le x_1$ and $y_2 \le x_2$ are also
compatible if $x_1$ and $x_2$ are compatible.

\begin{lemma}
\textit{noTransitWrite}: Whenever data for an address $a$ is in transit (\ie{}
$\forall T. \; T_s \le T \le T_r$ where $T_s$ is the time of sending the data and
$T_r$ the time of receiving the data), no cache can process a store request for
$a$, and the data must be sent from a \textit{clean} cache.
\label{noTransitWrite}
\end{lemma}

\begin{lemma}
\textit{conservative}: The state of an address in a cache is never greater than
the corresponding entry in its parent's directory.
\label{conservative}
\end{lemma}

\begin{lemma}
\textit{localCompatibility1}:
The state of a cache is never less than its directory entry for any of
its children.
\label{localCompatibility1}
\end{lemma}
\begin{lemma}
\textit{localCompatibility2}: Let $c_1$ and $c_2$ be distinct caches with the
same parent $p$. Then, $\s(c_1, a)$ and $\s(c_2, a)$ are
\textit{compatible}.
\label{localCompatibility2}
\end{lemma}

Lemmas~\ref{conservative} and~\ref{localCompatibility1} imply a
$\ge$ relation between the states of a parent and its child for any address.
Since the ancestor relation is the reflexive-transitive closure of the parent
relation, the $\ge$ relation is preserved between the states of two caches $p$
and $c$ where $p$ is the ancestor of $c$.

Any two caches $c_1, c_2$ neither of which is an ancestor of the other have a
lowest common ancestor $x$, which has
two distinct children $x_1$ and $x_2$ such that they are the ancestors of $c_1$
and $c_2$ respectively. For any address $a$, $\s(x_1, a)$ and
$\s(x_2, a)$ are compatible by Lemma~\ref{localCompatibility2}. From
the previous paragraph about the relation between the states of a cache and its
ancestor, it follows that $c_1$ and $c_2$ are also compatible, proving
Lemma~\ref{nonAncestorCompatible}.

Using Lemmas~\ref{nonAncestorCompatible} and~\ref{noTransitWrite}, one can
prove Lemma~\ref{latestValue}. The proof is by induction on time $T$; we assume
that any \clean{} (node, address) pair has the latest value until time
$T$. Consider a (node, addr) pair at time $T+1$.
If the pair was \clean{} at $T$, then no cache,
save itself (because of Lemma~\ref{nonAncestorCompatible}), can process any
store request for that address, thus continuing to be \clean{} and having the latest value.
And if the pair was not \clean{} at time $T$, it will receive data from a cache
that is \clean{} for the (node, address) pair, and no store
request can be processed when the data is in transit (by
Lemma~\ref{noTransitWrite}). Therefore, by the induction hypothesis the cache will
continue to have the latest value whenever it is \clean{} for an address at time $T+1$.

Analyzing a parent-child pair will suffice in proving
Lemmas~\ref{localCompatibility1} and \ref{localCompatibility2}.  Proof of
Lemma~\ref{noTransitWrite} is similar to the proof of
Lemma~\ref{nonAncestorCompatible}.

\section{The Final Result}\label{sec:finalresult}

With our two main results about optimized processors and memories, it
is now straightforward to complete the correctness proof of the
composed optimized system.

First, we need to know that, whenever the simple memory can generate
some trace of messages, it could also generate the same trace with all
speculative messages removed.  We need this property to justify the
introduction of speculation, during our final series of refinements
from the optimized system to SC.

\begin{theorem}
\label{mcorrect}
$M_m \sqsubseteq_{\noSpec^n} M_m$
\end{theorem}
\begin{proof}
By induction on traces, with an identity abstraction function.
\end{proof}

That theorem turns out to be the crucial ingredient to justify placing
a speculative processor in-context with simple memory.

\begin{theorem}
\label{ocorrect1}
$\text{P$^n_\text{so}$} + M_m \sqsubseteq \text{P$^n_\text{ref}$} + M_m$
\end{theorem}
\begin{proof}
Follows from Theorem \ref{liftplus} (one of our preliminary results
about $+$), Corollary \ref{ges}, and Theorem \ref{mcorrect}.
\end{proof}

The last theorem kept the memory the same while refining the
processor.  The next one does the opposite, switching out memory.

\begin{theorem}
\label{ocorrect2}
$\text{P$^n_\text{so}$} + M_c \sqsubseteq \text{P$^n_\text{so}$} + M_m$
\end{theorem}
\begin{proof}
Follows from Theorems \ref{ccorrect} and \ref{liftplus} plus reflexivity of $\sqsubseteq$ (Theorem \ref{transitive}).
\end{proof}

\begin{theorem}
\label{ofull}
$\text{P$^n_\text{so}$} + M_c \sqsubseteq SC$
\end{theorem}
\begin{proof}
We twice apply $\sqsubseteq$ transitivity (Theorem \ref{transitive}) to
%connect Theorems \ref{ocorrect2} and \ref{ocorrect1}, and use Definition \ref{sc} of $SC$
connect Theorems \ref{ocorrect2}, \ref{ocorrect1}, and \ref{scthm}.
\end{proof}