### 2014-11-09 RelatedWork.tex
stuff

\section{Related Work}
\label{relatedWork}

%Formal verification has been used to verify multiprocessor
%systems, especially with model checking
%\cite{burch1994automatic,mcmillan1998verification}.

%More recently, Intel's execution cluster (\ie{} the ALU and its
%interface) has been formally verified using model checking
%\cite{kaivola2009replacing}.

Theorem provers have also been used to verify microprocessors, \eg{} HOL has been
used to verify an academic microprocessor AVI-1 \cite{windley1995formal}.
Term-rewriting based techniques have been used to specify microprocessors, with
the goal of making them more amenable to verification \cite{shen1999using}.

Theorem provers have also been used in verifying cache coherence protocols.
The correctness of the FLASH cache coherence protocol~\cite{flash} for a
single-level hierarchy was proven with PVS~\cite{park} using a mapping, called
``aggregation of transactions,'' from multiple localized transactions to a
single global transaction, and finally using model checking to prove the correctness of
the global transactions model. However, they do not provide a methodology to
prove the aggregation of transactions for other protocols, which is where, we
believe, the difficulty lies; in contrast, our modular proof relies on generic
invariants, making it more adaptable to different protocols.
Cachet~\cite{StoyShenArvind:Proofs}, another single-level cache coherence
protocol, was verified to be obey the CRF memory
model~\cite{Shen:CRF}, and parts of the proof were implemented using PVS.

% Shen \etal{} provided a paper-and-pencil proof showing
%Cachet does not violate the CRF memory model, while Stoy \etal{} implemented
%portions of the paper-and-pencil proof using PVS. However, neither proof
%establishes that the protocol is store atomic. Rather, they show that Cachet
%obeys the CRF memory model, which is not restricted to store-atomic behavior.

There has been a lot of work in providing frameworks for axiomatic
specifications of memory models based on empirically observed behavior of
certain ``litmus
tests''\cite{Alglave11,Alglave:FMSD,alglave2012formal,Alglave:2010,Alglave:TACAS,alglave2014herding}.
There has been some work in providing operational and axiomatic semantics that
describes the observed behavior (based on certain ``litmus tests'' of real
hardware
\cite{mador2012axiomatic,sarkar2011understanding,sarkar2012synchronising}.
Finally, there has been work formally relating the axiomatic specification of
TSO to operational semantics \cite{x86tsocacm10}.

%There has been a lot of work in providing frameworks for axiomatic
%specifications of memory models \cite{alglave2012formal,alglave2014herding}
%and in providing axiomatic memory model specifications for real systems
%through analysis of empirically observed behavior of certain ``litmus tests''
%\cite{mador2012axiomatic,sarkar2012synchronising}. There has also been some
%work in providing operational semantics that describes the observed behavior of
%real hardware \cite{sarkar2011understanding}. Finally, there has been work
%formally relating the axiomatic specification of TSO to operational semantics\cite{x86tsocacm10}.

%The plethora of work in the area of memory models indeed suggests that this is
%a complex problem, exacerbated by the hardware architect's desire to increase
%hardware performance at the cost of complicating the memory models further. There
%are two aspects where our work can help in tackling this complex problem: (1) we
%are proposing a modular approach for verifying systems in which the components,
%namely the processors and the memory, can be verified against the memory model
%contract independent of each other, and (2) we are also proposing a design
%in which a system obeying sequential consistency is made into a high
%performance machine through the use of speculative loads followed by
%verification. The only downside is waiting for store acknowledgements, but this
%wait does not preclude the machine from performing further computations for
%future instructions without actually committing them. We are hoping to develop
%both these ideas -- of extending our proof approach for more complex memory
%models and designing real hardware which is sequentially consistent but has
%high performance.
%

%BELOW IS UNEDITTED FROM OLD PAPER

%These are lifted as is from that guy. Need to figure out how to restate them, and make it more concise.
%
%For verification of high level specifications, modern industrial practice
%consists of modeling small instances of the protocols, e.g., three CPUs
%handling two addresses and one bit of data, in terms of interleaving atomic
%steps, in guard/action languages such as Murphi [50] or TLA+ [84], and
%exploring the reachable states through explicit state enumeration.
%Symbolic methods, e.g., BDD [31] or SAT [129, 142], are yet to succeed for the
%verification of cache coherence protocols.
%Monolithic formal verification methods – methods that treat the protocol as a whole – have
%been used fairly routinely for verifying cache coherence protocols from the early 1990s
%[10 (2 nodes, murphi)]. However, these monolithic techniques will not be able to handle the very
%large state space of hierarchical protocols. Compositional verification techniques are essential. 
%
%For coherence protocols which only consider one level (or nonhierarchical protocols),
%various techniques have been proposed for the verification. These techniques basically
%include explicit state enumeration [133] and symbolic model checking [102]
%Opti-
%mizations on these techniques include partial order reduction [24], symmetry reduction
%[25, 72], compositional reasoning [96], predicate abstraction [22, 49, 82], etc. There is
%also a large body of work on parameterized verification [42, 53, 54, 100, 117]. However,
%none of them have shown the capability to verify hierarchical coherence protocols with
%realistic complexity.
%
%McMillan et al. [97, 102] modeled a 2-level MSI coherence protocol based on the
%Gigamax distributed multiprocessor [102]. In the protocol, bus snooping is used in both
%levels. SMV [97] was used to verify this protocol with two clusters each having six
%processors, for both safety and liveness properties. In terms of the complexity, many
%realistic details are abstracted away from the protocol so that the protocol does not have
%the typical complexity that a hierarchical coherence protocol has.
%
%=======
%Theorem provers have also been used in verifying cache coherence protocols,
%\eg{} FLASH~\cite{park} and Cachet~\cite{Shen:CRF, StoyShenArvind:Proofs} have
%been proved using the PVS theorem prover. Both works prove correctness of a
%single-level hierarchy while our work proves it for a multi-level hierarchy.
%The work by Park \etal{} on FLASH proves a mapping, called ``aggregation of
%transactions'', from multiple localized transactions to a single global
%transaction with PVS, and uses model-checking to prove the correctness of the
%global transactions model. However, they do not provide a methodology to prove
%the aggregation of transactions for other protocols, which we believe to be very
%difficult; in contrast, our modular proof relies on generic invariants making it
%more naturally adaptable to different protocols.
%>>>>>>> 96e9880a190d9acb6201b10e2b105d18c72f56aa