### 2014-11-10 Abstract.tex
Pass up to end of Introduction

We present a new framework for modular verification of hardware designs in the
style of the Bluespec language.  That is, we formalize the idea of components
in a hardware design, with well-defined input and output channels; and we show
how to specify and verify components individually, with machine-checked proofs
in the Coq proof assistant.  As a demonstration, we verify a fairly realistic
implementation of a multicore shared-memory system with two types of
components: memory system and processor.  Both components include nontrivial
optimizations, with the memory system employing an arbitrary hierarchy of cache
nodes that communicate with each other concurrently, and with the processor
doing speculative execution of many concurrent read operations.  Nonetheless,
we prove that the combined system implements sequential consistency.  One
interesting aspect of the proof is that the memory component is well-specified
to be reusable for systems implementing a variety of weak memory models.  To
our knowledge, our memory-system proof is the first machine verification
of a cache-coherence protocol parameterized over an arbitrary cache hierarchy,
and our full-system proof is the first machine verification of sequential
consistency for a multicore hardware design that includes caches and
speculative processors.