### 2014-12-14 Rebuttal.txt
added reviews

Some of the questions raised by the first three reviewers are addressed in the
comments of the Reviewer D.

We have submitted the source code of our Coq development as a suplement to our
paper in the submission website. It can also be obtained from the public repository
http://github.com/vmurali/SeqConsistency.

In terms of code-reuse, we have structured the proofs modularly as described in
the paper. Thus, the proofs of the different components can be reused while
changing the other components. For example, the proof of the cache-coherence
protocol can be reused as is for verifying other (non-sequentially consistent)
systems. Moreover, the proof of the cache-coherence is itself modular and can
be adopted to other protocols (like MESI, MOSI); The proofs for Lemmas 2 to 5
will change, but the lemmas correspond to at most 3 components (and states) and
hence can be easily proved (even using techniques like model-checking).

We agree with Reviewer-D that liveness is an important non-trivial property
that has to be proven for a cache-coherence protocol. The example of
ParentSendReq is an example of a voluntary rule, and the policy to fire such a
rule does indeed affect the liveness. In our development, we haven't proved
full liveness (which includes the liveness of the processor), but we have
proved the following (stated informally):
a) Any request (or response) that a cache wants to send can eventually be sent
(either to its parent or to its children)
b) Under the assumption that ParentSendReq rule (sending downgrade request from
a parent to its children) will fire eventually, whenever the parent has an
upgrade request from a child, every request will eventually get a response.

Thus, the only extraneous condition we impose is that a parent should "decide"
to send a downgrade request to (some of) its children whenever it sees an
upgrade request from one of its children. Eventually, such a request will be
sent via ParentSendReq rule.

We believe such a modular technique for verifying full-system correctness with
respect to SC will be useful for hardware designers/verification-engineers to
independently verify different components of the system. We do not model non-SC
systems (like ARM/POWER), which exhibit behaviors described by Reviewer-C
(involving shared store buffers, committing non-verified load values, etc).
Reviewer-C, please note that the idea of verification loads does indeed ensure
correctness, even if there were stores to that address in between the
speculative load and the verification load, as long as the final store read by
the verification load writes the same value as the speculative load. In fact,
our final theorem ensures that this is indeed the case, otherwise the
speculative system will not be an implementation of the SC reference system.

Reviewer-B please note that our SC model is exactly as specified by Lamport.
Also please note that by local buffers in Figure 2, we do not mean just store
buffers; it includes all local data accessible to the processor.

Thanks for notifying us of other related work (Hu et.al.'s "Decidable SC"
SPAA'03, Qadeer's "Verifying SC by model-checking" TPDS'03, Sorin's
"Verification-aware Architectures" PACT'07, "Lamport Clocks: Verifying
Directory Cache-Coherence Protocol" SPAA'98 and Moore's "ACL2 proof of cache
coherence" CAV'98), and for pointing the other bugs in our related work
section.

Reviewer-B's concerns:
- Ideal processor and memory: as described in that paragraph - memory responds
  instantaneously and processor executes atomically without speculation
- Bluespec served as the model: The motivation for LTS comes from using
  Bluespec to describe hardware, which is similar to our LTS formalization.
- Bluespec's competitive performance can be seen from [Dave et al. 2005], and
  other papers using Bluespec designs.
- Peek-beneath abstraction boundaries: To prove the full-system, the
  individual components are treated as black-boxes, completely captured by their
  interface specification.
- Good-ideas from worlds of PL semantics: Using LTSes as a precise hardware
  specification.
- Infinite families of hardware design ... resist reduction to finite-state
  models: reduction is applied to 'infinite families'
- Figure 2: logical view of system as 3-components: It is more-or-less accepted
  in the architecture community.
- Refining vs Equivalent: Throughout the paper, we are showing only refinement
  in one direction, though you are correct that in this particular example they
  are equivalent.
- Finer-grained architectural details: as mentioned in that paragraph (like
  register-renaming)
- Main-memory still exposes atomic loads and stores: Most architectures (including
  POWER and ARM) can be modeled using a memory that responds to loads and stores
  instantaneously, though the response (or request) can sit in a buffer between
  the memory and the processor, leading to relaxed memory models.