### 2015-01-21 Rebuttal2.txt
Clean version

One of the major benefits of our approach is that it unifies the modularization
of the design and its verification task which enables verification to be done
in tandem with mirco-architectural design (as noted by Reviewer-E). Further,
this allows us to require only incremental verification effort for an
incremental microarchitectural change, enabling scaling of the verification
task.

Our approach vs. Model-Checking-based:

a) Deriving Invariants:
Finding invariants is fundamentally hard. The CMP-method (referred by
Reviewer-E) uses symmetry to abstract the system, and then inserts invariants
(called non-interference-lemmas) manually to eliminate counter-examples that
violate the required property (the flow-based methodology [1] specifies these
non-interference-lemmas in the form of properties of action sequences that
takes place for any high-level transaction.) All these invariants are
relatively simple but very subtle and are well understood by the designers. For
example, for the FLASH protocol, one of the non-interference-lemma (Lemma-5 in
[2]) says that whenever a cache is in E state, then the data it has is
up-to-date. This is very similar to our Lemma-1. In general, we believe that
the level of complexity of the lemmas in the two approaches are similar.

The verification of the hierarchical protocol in [3] is restricted to 2-levels
whereas our work applies to arbitrary levels of hierarchy. We are not sure if
it's a fundamental restriction on the CMP-method, since the kind of symmetry
that CMP-method exploits in a one or two-level system is not present in a
hierarchical system.

b) Case-Analysis:
Much of the requisite protocol case analysis is done automatically via
proof-script automation in Coq. In our experience, cases where proof-script
automation fail required an additional invariant; we believe that manually
providing a similiar invariant is necessary even in a model-checking approach.

Reviewer-F:

Figure-3, no tags in Figures-5,6:
Both these points are valid and we should have indicated them explicitly.

Figure-6, third case:
The "upd" function merely determines if the "real" state of the system is as
dictated by (s,pc) or if it should be modified to take into account memory
responses. If the processor is not waiting for a memory response, then the
state is exactly (s,pc). We made an optimization in the fourth case where we
make the real state H whenever we know that the next instruction we decode is
H; we could have it as (s,pc) instead.

Figure-8: tags for non-speculative loads:
Our invariant is that there can be only one outstanding non-speculative load,
though there can be several outstanding speculative loads concurrently.
Anything without a tag is the current non-speculative load, and anything with a
tag is a speculative load.

Reviewer-E:

Pipelining protocol-engines:
While atomic transitions in our paper have one-at-a-time semantics, the
Bluespec compiler generates hardware which executes independant transitions
concurrently in a clock cycle (as the semantics aren't violated); the compiler
excels at this optimization. Atomic transition can also be spread across
multiple clock cycles, while maintaining the semantics [4].

Evidence of machine-checked proofs:
We submitted our Coq source code as a supplement to our paper in the submission
website.

On Describing LTSes:
We include the LTS description for the benefit of readers unfamiliar with it;
the point wasn't to present novel concepts.

References:
[1] Parameterized Verification using Message Flows: An Industrial Experience,
John O'Leary, Murali Talupur, Mark R. Tuttle, FMCAD'09
[2] A Simple Method for Parameterized Verification of Cache Coherence Protocols,
Ching-Tsun Chou, Phanindra K. Mannava, Seungjoon Park, FMCAD'04
[3] Efficient methods for formally verifying safety properties of hierarchical
cache coherence protocol, Xiaofang Chen, Yu Yang, Ganesh Gopalakrishnan,
Ching-Tsun Chou, FMSD'10
[4] A new synthesis procedure for atomic rules containing multi-cycle function
blocks, Michal Karczmarek, Arvind, Muralidaran Vijayaraghavan, MEMOCODE'14